
```{r}
temp <- df |>
  add_all_firm_data() |> 
  mutate(value_final = as.numeric(value_final))

```

### Environmental performance early adopters vs. all

```{r}
# Median plots early adopters vs all
#| fig-width: 15
#| fig-height: 16

selected_indicators <- c("Total scope 1 emissions [tCO₂eq]",
                         "Total scope 2 emissions (market-based) [tCO₂eq]",
                         "Total scope 3 emissions [tCO₂eq]",
                         #"Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]",
                         "Category 6 Travel [tCO₂eq]",
                         #"Scope 3 emissions (Commuting) [tCO₂eq]",
                         "Energy consumption [MWh]",
                         "Renewable sources in total energy consumption [%]",
                         #"Fossil sources in total energy consumption [%]",
                         "Water consumption [m³]",
                         "Total waste generated [t]",
                         "Non-recycled waste [%]"
                        )
# define bins 
n_breaks <- c(0, 150, 300, Inf)
n_labels <- c("≤150", "≤300", ">300")

temp2 <- temp |>
  mutate(label_specification = recode(label_specification, !!!renamer_env),
         label_specification = factor(label_specification, levels = unname(renamer_env))
          ) |>
  select(firm, year, label_specification, value_final) |>
  filter(label_specification %in% selected_indicators, !is.na(value_final))

# Get early adopters
early_adop_firms <- temp2 |>
  filter(year %in% c(2014, 2015)) |>
  distinct(firm, label_specification)

temp2 <- temp2 |>
  left_join(
    early_adop_firms |> mutate(early_adopter = TRUE),
    by = c("firm", "label_specification")
  ) |>
  mutate(early_adopter = replace_na(early_adopter, FALSE))


grouping_info <- temp2 |>
  filter(early_adopter == TRUE) |>  # focus only on early adopters
  group_by(label_specification, year) |>
  summarise(
    n = n(),
    .groups = "drop"
  ) |>
  mutate(
    n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
    n_group = factor(n_group, levels = n_labels)
  )

p_all <- temp2 |>
  select(firm, year, label_specification, value_final) |>
  group_by(label_specification, year) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    early_adopter = "All companies",
    .groups = "drop"
  )

p_early <- temp2 |>
  filter(early_adopter == TRUE) |>
  group_by(label_specification, year) |>
  mutate(
    lower_bound = quantile(value_final, 0.00, na.rm = TRUE),
    upper_bound = quantile(value_final, 0.99, na.rm = TRUE),
    value_final = pmin(pmax(value_final, lower_bound), upper_bound)
  ) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    early_adopter = "Early adopters",
    .groups = "drop"
  )

p_env <- bind_rows(p_all, p_early) |>
  left_join(grouping_info, by = c("label_specification", "year")) |>
  mutate(
    n_group = factor(n_group, levels = n_labels),
    early_adopter = factor(early_adopter, levels = c("All companies", "Early adopters"))
  )

n_values <- temp2 |>
  filter(early_adopter == TRUE) |>
  group_by(label_specification) |>
  summarise(n = n(), .groups = "drop")

# set y axis limits independently for emission metrics
emission_indicators <- c(
  "Total Scope 1 emissions [tCO₂eq]",
  "Total scope 2 emissions (market-based) [tCO₂eq]",
  "Total scope 3 emissions [tCO₂eq]",
  "Category 6 Travel [tCO₂eq]"
)

shared_emission_max <- p_env |>
  filter(label_specification %in% emission_indicators) |>
  summarise(ymax = max(p90, q3, na.rm = TRUE)) |>
  pull(ymax) * 1.1


plotting <- function(indicator) {
  indicator_title <- switch(indicator,
    "Total scope 1 emissions [tCO₂eq]" = expression(Total~scope~1~emissions~"[tCO"[2]*"eq]"),
    "Total scope 2 emissions (market-based) [tCO₂eq]" = expression(Total~scope~2~emissions~"[tCO"[2]*"eq]"),
    "Total scope 3 emissions [tCO₂eq]" = expression(Total~Scope~3~emissions~"[tCO"[2]*"eq]"),
    "Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]" = expression(Scope~3~emissions~(Cloud~computing~and~data~center~services)~"[tCO"[2]*"eq"),
    "Category 6 Travel [tCO₂eq]" = expression(Scope~3~emissions~(Travel)~"[tCO"[2]*"eq]"),
    "Scope 3 emissions (Commuting) [tCO₂eq]" = expression(Scope~3~emissions~(Commuting)~"[tCO"[2]*"eq"),
    "Energy consumption [MWh]" = expression(Energy~consumption~"[MWh]"),
    "Renewable sources in total energy consumption [%]" = expression(atop("Renewable sources in total", "energy consumption [%]")),
    "Fossil sources in total energy consumption [%]" = expression(atop("Fossil sources in total", "energy consumption [%]")),
    "Water consumption [m³]" = expression(Water~consumption~"[m"^3*"]"),
    "Total waste generated [t]" = expression(Total~waste~generated~"[t]"),
    "Non-recycled waste [%]" = expression("Non-recycled waste [%]"),
    expression(.(indicator))
  )
 data_filtered <- p_env |>
  filter(label_specification == indicator) |>
  mutate(n_group = factor(n_group, levels = n_labels)) |>
  complete(
    year,
    label_specification,
    early_adopter = c("All companies", "Early adopters"),
    n_group = factor(n_labels, levels = n_labels)
  )

  n_label <- n_values |>
    filter(label_specification == indicator) |>
    pull(n)

  label_n <- paste0("italic(n)==", n_label)

  # use dynamic y-axis limits
  y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
  y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

  y_max <- if (indicator %in% emission_indicators) {
    shared_emission_max
  } else {
    max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1
  }

  p <- ggplot(data_filtered, aes(x = year))

  if (indicator == "Scope 3 emissions (Travel) [tCO₂eq]") {
      # add shaded area in the background
      p <- p +
        geom_rect(
          data = tibble(xmin = 2020, xmax = 2023, ymin = 1, ymax = 12000000),
          aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
          fill = "grey85", alpha = 0.3, inherit.aes = FALSE
        ) +

        # add annotation text in the foreground
        annotate(
          "text",
          x = 2020.3, y = 500000,
          label = "COVID-19\npandemic",
          hjust = 0, size = 4, color = "grey30"
        )
  }

  p <- p +
    geom_line(data = filter(data_filtered, !is.na(median)),
              aes(y = median, color = early_adopter, linetype = early_adopter),
              size = 1) 
    # geom_point(
    #   data = filter(data_filtered, early_adopter == "Early adopters"),
    #   aes(y = median, shape = n_group, size = n_group, fill = n_group),
    #   color = "#2E6F40"
    # ) +
    if (indicator %in% c(
    "Renewable sources in total energy consumption [%]",
    "Non-recycled waste [%]"
    )) {
      p <- p +
        scale_y_continuous(
          limits = c(
            min(data_filtered$p10, na.rm = TRUE) * 0.9,
            max(data_filtered$p90, na.rm = TRUE) * 1.1
            ),
          labels = scales::comma_format()
          )
      } else {
        p <- p +
          scale_y_continuous(
            trans = "log10",
            labels = scales::comma_format(),
            limits = c(1, y_max),
            oob = scales::oob_keep
          )
      }
  
  p <- p +
    scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
    scale_color_manual(
      values = c("All companies" = "grey", "Early adopters" = "#2E6F40"),
      name = NULL
    ) +
    scale_linetype_manual(
      values = c("All companies" = "dashed", "Early adopters" = "solid"),
      name = NULL
    ) +

    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
      panel.grid.minor.y = element_blank(),
      axis.ticks = element_line(color = "grey"),
      axis.ticks.length = unit(0.15, "cm"),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16)
    ) +
    labs(
      title = bquote(.(indicator_title)),
      x = "",
      y = parse(text = "")
    )

  label_grob <- grid::textGrob(
      label = parse(text = label_n),
      x = unit(0.98, "npc"),
      y = unit(0.95, "npc"),
      just = c("right", "bottom"),
      gp = gpar(fontsize = 14)
      )

    p <- p + annotation_custom(label_grob)

    ggsave(
      filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
      plot = p,
      width = 5, height = 5, dpi = 400, units = "in", device = 'png'
    )

    return(p)
}


plots <- lapply(selected_indicators, plotting)

pfinal <- wrap_plots(plots) +
  plot_layout(
    ncol = 3,
    guides = "collect"
  ) +
  plot_annotation(
    tag_levels = "a"
  ) &
  theme(
    plot.tag = element_text(face = "bold", size = 20),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.justification = "center",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 16)
  )

ggsave(
    here("fig", "fig-env-impacts-early-adop.pdf"),
    pfinal,
    width = 16, height = 15, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-env-impacts-early-adop.pdf"))

```

### Social performance early adopters vs. all

```{r}
#| fig-width: 15
#| fig-height: 16

selected_indicators <- c("Employee turnover [%]",
                         "Employees covered by collective bargaining agreements [%]",
                         "# Training hours per employee",
                         "# Days lost to work-related injuries, ill health, and fatalities among employees",
                         #"Employees entitled to take family-related leaves [%]",
                         "Female employees in top managment [%]",
                         "Gender pay gap [%]",
                         "Annual remuneration ratio [%]",
                         "# Complaints filed by own workforce",
                         "Amount of fines, penalties, and compensation for damages as result of incidents and complaints [USD]"
                         #"# Severe human rights incidents connected to own workforce"
                        )

# define bins 
n_breaks <- c(0, 150, 300, Inf)
n_labels <- c("≤150", "≤300", ">300")

temp2 <- temp |>
  mutate(label_specification = recode(label_specification, !!!renamer_soc),
         label_specification = factor(label_specification, levels = unname(renamer_soc))
          ) |>
  select(firm, year, label_specification, value_final) |>
  filter(label_specification %in% selected_indicators, !is.na(value_final))

# Get early adopters
early_adop_firms <- temp2 |>
  filter(year %in% c(2014, 2015)) |>
  distinct(firm, label_specification)

temp2 <- temp2 |>
  left_join(
    early_adop_firms |> mutate(early_adopter = TRUE),
    by = c("firm", "label_specification")
  ) |>
  mutate(early_adopter = replace_na(early_adopter, FALSE))


grouping_info <- temp2 |>
  filter(early_adopter == TRUE) |>  # focus only on early adopters
  group_by(label_specification, year) |>
  summarise(
    n = n(),
    .groups = "drop"
  ) |>
  mutate(
    n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
    n_group = factor(n_group, levels = n_labels)
  )

p_all <- temp2 |>
  select(firm, year, label_specification, value_final) |>
  group_by(label_specification, year) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10 = if_else(
      label_specification != "Gender pay gap [%]",
      pmax(quantile(value_final, 0.10, na.rm = TRUE), 1),
      quantile(value_final, 0.10, na.rm = TRUE)
    ),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    early_adopter = "All companies",
    .groups = "drop"
  )

p_early <- temp2 |>
  filter(early_adopter == TRUE) |>
  group_by(label_specification, year) |>
  mutate(
    lower_bound = quantile(value_final, 0.00, na.rm = TRUE),
    upper_bound = quantile(value_final, 0.99, na.rm = TRUE),
    value_final = pmin(pmax(value_final, lower_bound), upper_bound)
  ) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10 = if_else(
      label_specification != "Gender pay gap [%]",
      pmax(quantile(value_final, 0.10, na.rm = TRUE), 1),
      quantile(value_final, 0.10, na.rm = TRUE)
    ),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    early_adopter = "Early adopters",
    .groups = "drop"
  )

p_soc <- bind_rows(p_all, p_early) |>
  left_join(grouping_info, by = c("label_specification", "year")) |>
  mutate(
    n_group = factor(n_group, levels = n_labels),
    early_adopter = factor(early_adopter, levels = c("All companies", "Early adopters"))
  )

n_values <- temp2 |>
  filter(early_adopter == TRUE) |>
  group_by(label_specification) |>
  summarise(n = n(), .groups = "drop")


plotting <- function(indicator) {
  
 data_filtered <- p_soc |>
  filter(label_specification == indicator) |>
  mutate(n_group = factor(n_group, levels = n_labels)) |>
  complete(
    year,
    label_specification,
    early_adopter = c("All companies", "Early adopters"),
    n_group = factor(n_labels, levels = n_labels)
  )

  n_label <- n_values |>
    filter(label_specification == indicator) |>
    pull(n)

  label_n <- paste0("italic(n)==", n_label)

  # use dynamic y-axis limits
  y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
  y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

  y_max <- max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1

  p <- ggplot(data_filtered, aes(x = year)) +
    
    geom_line(data = filter(data_filtered, !is.na(median)),
              aes(y = median, color = early_adopter, linetype = early_adopter),
              size = 1)
    # geom_point(
    #   data = filter(data_filtered, early_adopter == "Early adopters"),
    #   aes(y = median, shape = n_group, size = n_group, fill = n_group),
    #   color = "#2E6F40"
    # ) +
    if (indicator %in% c("Employee turnover [%]",
                         "Employees covered by collective bargaining agreements [%]",
                         "Female employees in top managment [%]",
                         "Gender pay gap [%]",
                         "Annual remuneration ratio [%]")) {
      p <- p + scale_y_continuous(
        limits = c(min(data_filtered$p10, na.rm = TRUE) * 0.9,
                   max(data_filtered$p90, na.rm = TRUE) * 1.1),
        labels = scales::comma_format()
        )
    } else {
        p <- p + scale_y_continuous(
          trans = "log10",
          labels = scales::comma_format(),
          limits = c(1, y_max),
          oob = scales::oob_keep
          )
    }
  
  p <- p +
    
    scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
    scale_color_manual(
      values = c("All companies" = "grey", "Early adopters" = "#6F2E5D"),
      name = NULL
    ) +
    scale_linetype_manual(
      values = c("All companies" = "dashed", "Early adopters" = "solid"),
      name = NULL
    ) +

    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
      panel.grid.minor.y = element_blank(),
      axis.ticks = element_line(color = "grey"),
      axis.ticks.length = unit(0.15, "cm"),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16)
    ) +
    labs(
        title = stringr::str_wrap(indicator, width = 45),
        x = "",
        y = parse(text = "")
    )

  label_grob <- grid::textGrob(
      label = parse(text = label_n),
      x = unit(0.98, "npc"),
      y = unit(0.95, "npc"),
      just = c("right", "bottom"),
      gp = gpar(fontsize = 14)
      )

    p <- p + annotation_custom(label_grob)

    ggsave(
      filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
      plot = p,
      width = 5, height = 5, dpi = 400, units = "in", device = 'png'
    )

    return(p)
}


plots <- lapply(selected_indicators, plotting)

pfinal <- wrap_plots(plots) +
  plot_layout(
    ncol = 3,
    guides = "collect"
  ) +
  plot_annotation(
    tag_levels = "a"
  ) &
  theme(
    plot.tag = element_text(face = "bold", size = 20),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.justification = "center",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 16)
  )

ggsave(
    here("fig", "fig-soc-impacts-early-adop.pdf"),
    pfinal,
    width = 16, height = 17, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-soc-impacts-early-adop.pdf"))


```

### Governance performance early adopters vs. all

```{r}
#| fig-width: 15
#| fig-height: 5.5

selected_indicators <- c("Independent board members [%]",
                         #"# Confirmed incidents of corruption and bribery",
                         "Lobbying expenses [USD]",
                         #"Political contributions made [EUR]",
                         #"Variable remuneration dependent on sustainability-targets [%]"
                         "# Days to pay invoice"
                        )
# define bins 
n_breaks <- c(0, 150, 300, Inf)
n_labels <- c("≤150", "≤300", ">300")

temp2 <- temp |>
  mutate(label_specification = recode(label_specification, !!!renamer_gov),
         label_specification = factor(label_specification, levels = unname(renamer_gov))
          ) |>
  select(firm, year, label_specification, value_final) |>
  filter(label_specification %in% selected_indicators, !is.na(value_final))

# Get early adopters
early_adop_firms <- temp2 |>
  filter(year %in% c(2014, 2015)) |>
  distinct(firm, label_specification)

temp2 <- temp2 |>
  left_join(
    early_adop_firms |> mutate(early_adopter = TRUE),
    by = c("firm", "label_specification")
  ) |>
  mutate(early_adopter = replace_na(early_adopter, FALSE))


grouping_info <- temp2 |>
  filter(early_adopter == TRUE) |>  # focus only on early adopters
  group_by(label_specification, year) |>
  summarise(
    n = n(),
    .groups = "drop"
  ) |>
  mutate(
    n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
    n_group = factor(n_group, levels = n_labels)
  )

p_all <- temp2 |>
  select(firm, year, label_specification, value_final) |>
  group_by(label_specification, year) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    early_adopter = "All companies",
    .groups = "drop"
  )

p_early <- temp2 |>
  filter(early_adopter == TRUE) |>
  group_by(label_specification, year) |>
  mutate(
    lower_bound = quantile(value_final, 0.00, na.rm = TRUE),
    upper_bound = quantile(value_final, 0.99, na.rm = TRUE),
    value_final = pmin(pmax(value_final, lower_bound), upper_bound)
  ) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    early_adopter = "Early adopters",
    .groups = "drop"
  )

p_gov <- bind_rows(p_all, p_early) |>
  left_join(grouping_info, by = c("label_specification", "year")) |>
  mutate(
    n_group = factor(n_group, levels = n_labels),
    early_adopter = factor(early_adopter, levels = c("All companies", "Early adopters"))
  )

n_values <- temp2 |>
  filter(early_adopter == TRUE) |>
  group_by(label_specification) |>
  summarise(n = n(), .groups = "drop")

plotting <- function(indicator) {
  
 data_filtered <- p_gov |>
  filter(label_specification == indicator) |>
  mutate(n_group = factor(n_group, levels = n_labels)) |>
  complete(
    year,
    label_specification,
    early_adopter = c("All companies", "Early adopters"),
    n_group = factor(n_labels, levels = n_labels)
  )

  n_label <- n_values |>
    filter(label_specification == indicator) |>
    pull(n)

  label_n <- paste0("italic(n)==", n_label)

  # use dynamic y-axis limits
  y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
  y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

  y_max <- max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1

  p <- ggplot(data_filtered, aes(x = year)) +
    
    geom_line(data = filter(data_filtered, !is.na(median)),
              aes(y = median, color = early_adopter, linetype = early_adopter),
              size = 1)
    # geom_point(
    #   data = filter(data_filtered, early_adopter == "Early adopters"),
    #   aes(y = median, shape = n_group, size = n_group, fill = n_group),
    #   color = "#2E6F40"
    # ) +
     if (indicator %in% c("Independent board members [%]")) {
      p <- p + scale_y_continuous(
        limits = c(min(data_filtered$p10, na.rm = TRUE) * 0.9,
                   max(data_filtered$p90, na.rm = TRUE) * 1.1),
        labels = scales::comma_format()
        )
    } else {
        p <- p + scale_y_continuous(
          trans = "log10",
          labels = scales::comma_format(),
          limits = c(1, y_max),
          oob = scales::oob_keep
          )
    }
  
  p <- p +
    scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
    scale_color_manual(
      values = c("All companies" = "grey", "Early adopters" = "#485DA5"),
      name = NULL
    ) +
    scale_linetype_manual(
      values = c("All companies" = "dashed", "Early adopters" = "solid"),
      name = NULL
    ) +

    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
      panel.grid.minor.y = element_blank(),
      axis.ticks = element_line(color = "grey"),
      axis.ticks.length = unit(0.15, "cm"),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16)
    ) +
    labs(
        title = stringr::str_wrap(indicator, width = 45),
        x = "",
        y = parse(text = "")
    )

  label_grob <- grid::textGrob(
      label = parse(text = label_n),
      x = unit(0.98, "npc"),
      y = unit(0.95, "npc"),
      just = c("right", "bottom"),
      gp = gpar(fontsize = 14)
      )

    p <- p + annotation_custom(label_grob)

    ggsave(
      filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
      plot = p,
      width = 5, height = 5, dpi = 400, units = "in", device = 'png'
    )

    return(p)
}


plots <- lapply(selected_indicators, plotting)

pfinal <- wrap_plots(plots) +
  plot_layout(
    ncol = 3,
    guides = "collect"
  ) +
  plot_annotation(
    tag_levels = "a"
  ) &
  theme(
    plot.tag = element_text(face = "bold", size = 20),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.justification = "center",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 16)
  )

ggsave(
    here("fig", "fig-gov-impacts-early-adop.pdf"),
    pfinal,
    width = 16, height = 6, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-gov-impacts-early-adop.pdf"))

```
### Environmental performance constant reporters vs. all

```{r}
#| fig-width: 15
#| fig-height: 16

library(jsonlite) # TEMP: append to analyses.qmd and requirements.txt

constant_reporters <- jsonlite::fromJSON(
  here::here("data/raw/datasets/constant_reporters.json")
) # TEMP: decide on where JSON file should be stored

selected_indicators <- c(
    "Total scope 1 emissions [tCO₂eq]",
    "Total scope 2 emissions (market-based) [tCO₂eq]",
    "Total scope 3 emissions [tCO₂eq]",
    # "Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]",
    "Category 6 Travel [tCO₂eq]",
    # "Scope 3 emissions (Commuting) [tCO₂eq]",
    "Energy consumption [MWh]",
    "Renewable sources in total energy consumption [%]",
    # "Fossil sources in total energy consumption [%]",
    "Water consumption [m³]",
    "Total waste generated [t]",
    "Non-recycled waste [%]"
)

# define bins
n_breaks <- c(0, 150, 300, Inf)
n_labels <- c("≤150", "≤300", ">300")

# prepare base data with renamed labels
temp_renamed <- temp |>
    select(firm, year, topic, id, label_specification, value_final) |>
    mutate(
        label_specification = recode(label_specification, !!!renamer_env),
        label_specification = factor(label_specification, levels = unname(renamer_env))
    ) |>
    filter(topic == "env", label_specification %in% selected_indicators) |>
    drop_na(value_final)

# create mapping from display name to indicator ID
tmp <- temp_renamed |>
  dplyr::distinct(id, label_specification)

name_to_id_map <- stats::setNames(tmp$id, as.character(tmp$label_specification))

# flag constant reporters for each observation
temp_flagged <- temp_renamed |>
    rowwise() |>
    mutate(
        ind_id = name_to_id_map[as.character(label_specification)],
        is_constant_reporter = if (!is.na(ind_id) && ind_id %in% names(constant_reporters)) {
            firm %in% constant_reporters[[ind_id]]
        } else {
            FALSE
        }
    ) |>
    ungroup() |>
    select(-ind_id)

# compute grouping info for constant reporters
grouping_info <- temp_flagged |>
    filter(is_constant_reporter == TRUE) |>
    group_by(label_specification, year) |>
    summarise(
        n = n(),
        .groups = "drop"
    ) |>
    mutate(
        n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
        n_group = factor(n_group, levels = n_labels)
    )

# compute summaries for ALL companies
p_all <- temp_flagged |>
    group_by(label_specification, year) |>
    summarise(
        p10 = quantile(value_final, 0.10, na.rm = TRUE),
        q1 = quantile(value_final, 0.25, na.rm = TRUE),
        median = quantile(value_final, 0.50, na.rm = TRUE),
        q3 = quantile(value_final, 0.75, na.rm = TRUE),
        p90 = quantile(value_final, 0.90, na.rm = TRUE),
        group = "All companies",
        .groups = "drop"
    )

# compute summaries for CONSTANT REPORTERS only
p_constant <- temp_flagged |>
    filter(is_constant_reporter == TRUE) |>
    group_by(label_specification, year) |>
    summarise(
        p10 = quantile(value_final, 0.10, na.rm = TRUE),
        q1 = quantile(value_final, 0.25, na.rm = TRUE),
        median = quantile(value_final, 0.50, na.rm = TRUE),
        q3 = quantile(value_final, 0.75, na.rm = TRUE),
        p90 = quantile(value_final, 0.90, na.rm = TRUE),
        group = "Constant reporters",
        .groups = "drop"
    )

# combine both
p_env <- bind_rows(p_all, p_constant) |>
    left_join(grouping_info, by = c("label_specification", "year")) |>
    mutate(
        n_group = factor(n_group, levels = n_labels),
        group = factor(group, levels = c("All companies", "Constant reporters"))
    )

# compute n values for constant reporters
n_values <- temp_flagged |>
    filter(is_constant_reporter == TRUE) |>
    group_by(label_specification) |>
    summarise(n = n(), .groups = "drop")

# set y axis limits independently for emission metrics
emission_indicators <- c(
    "Total Scope 1 emissions [tCO₂eq]",
    "Total scope 2 emissions (market-based) [tCO₂eq]",
    "Total scope 3 emissions [tCO₂eq]",
    "Category 6 Travel [tCO₂eq]"
)

shared_emission_max <- p_env |>
    filter(label_specification %in% emission_indicators) |>
    summarise(ymax = max(p90, q3, na.rm = TRUE)) |>
    pull(ymax) * 1.1


plotting <- function(indicator) {
    indicator_title <- switch(indicator,
        "Total scope 1 emissions [tCO₂eq]" = expression(Total ~ scope ~ 1 ~ emissions ~ "[tCO"[2] * "eq]"),
        "Total scope 2 emissions (market-based) [tCO₂eq]" = expression(Total ~ scope ~ 2 ~ emissions ~ "[tCO"[2] * "eq]"),
        "Total scope 3 emissions [tCO₂eq]" = expression(Total ~ Scope ~ 3 ~ emissions ~ "[tCO"[2] * "eq]"),
        "Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]" = expression(Scope ~ 3 ~ emissions ~ (Cloud ~ computing ~ and ~ data ~ center ~ services) ~ "[tCO"[2] * "eq"),
        "Category 6 Travel [tCO₂eq]" = expression(Scope ~ 3 ~ emissions ~ (Travel) ~ "[tCO"[2] * "eq]"),
        "Scope 3 emissions (Commuting) [tCO₂eq]" = expression(Scope ~ 3 ~ emissions ~ (Commuting) ~ "[tCO"[2] * "eq"),
        "Energy consumption [MWh]" = expression(Energy ~ consumption ~ "[MWh]"),
        "Renewable sources in total energy consumption [%]" = expression(atop("Renewable sources in total", "energy consumption [%]")),
        "Fossil sources in total energy consumption [%]" = expression(atop("Fossil sources in total", "energy consumption [%]")),
        "Water consumption [m³]" = expression(Water ~ consumption ~ "[m"^3 * "]"),
        "Total waste generated [t]" = expression(Total ~ waste ~ generated ~ "[t]"),
        "Non-recycled waste [%]" = expression("Non-recycled waste [%]"),
        expression(.(indicator))
    )

    data_filtered <- p_env |>
        filter(label_specification == indicator) |>
        mutate(n_group = factor(n_group, levels = n_labels)) |>
        complete(
            year,
            label_specification,
            group = c("All companies", "Constant reporters"),
            n_group = factor(n_labels, levels = n_labels)
        )

    n_label <- n_values |>
        filter(label_specification == indicator) |>
        pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1

    y_max <- if (indicator %in% emission_indicators) {
        shared_emission_max
    } else {
        max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1
    }

    p <- ggplot(data_filtered, aes(x = year))

    if (indicator == "Category 6 Travel [tCO₂eq]") {
        p <- p +
            geom_rect(
                data = tibble(xmin = 2020, xmax = 2023, ymin = 1, ymax = 12000000),
                aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
                fill = "grey85", alpha = 0.3, inherit.aes = FALSE
            ) +
            annotate(
                "text",
                x = 2020.3, y = 500000,
                label = "COVID-19\npandemic",
                hjust = 0, size = 4, color = "grey30"
            )
    }

    p <- p +
        geom_line(
            data = filter(data_filtered, !is.na(median)),
            aes(y = median, color = group, linetype = group),
            size = 1
        )

    if (indicator %in% c(
        "Renewable sources in total energy consumption [%]",
        "Non-recycled waste [%]"
    )) {
        p <- p +
            scale_y_continuous(
                limits = c(
                    min(data_filtered$p10, na.rm = TRUE) * 0.9,
                    max(data_filtered$p90, na.rm = TRUE) * 1.1
                ),
                labels = scales::comma_format()
            )
    } else {
        p <- p +
            scale_y_continuous(
                trans = "log10",
                labels = scales::comma_format(),
                limits = c(1, y_max),
                oob = scales::oob_keep
            )
    }

    p <- p +
        scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
        scale_color_manual(
            values = c("All companies" = "grey50", "Constant reporters" = "#2E6F40"),
            name = NULL
        ) +
        scale_linetype_manual(
            values = c("All companies" = "dashed", "Constant reporters" = "solid"),
            name = NULL
        ) +
        theme_minimal() +
        theme(
            plot.title = element_text(size = 14, hjust = 0.5),
            panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
            panel.grid.major.x = element_blank(),
            panel.grid.minor.x = element_blank(),
            panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
            panel.grid.minor.y = element_blank(),
            axis.ticks = element_line(color = "grey"),
            axis.ticks.length = unit(0.15, "cm"),
            axis.text.x = element_text(size = 16),
            axis.text.y = element_text(size = 16)
        ) +
        labs(
            title = bquote(.(indicator_title)),
            x = "",
            y = parse(text = "")
        )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
    )

    p <- p + annotation_custom(label_grob)

    ggsave(
        filename = here("fig", paste0("fig-constant-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = "png"
    )

    return(p)
}


plots <- lapply(selected_indicators, plotting)

pfinal <- wrap_plots(plots) +
    plot_layout(
        ncol = 3,
        guides = "collect"
    ) +
    plot_annotation(
        tag_levels = "a"
    ) &
    theme(
        plot.tag = element_text(face = "bold", size = 20),
        legend.position = "bottom",
        legend.box = "vertical",
        legend.justification = "center",
        legend.title = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 16)
    )

ggsave(
    here("fig", "fig-env-impacts-constant-reporters.pdf"),
    pfinal,
    width = 16, height = 15, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-env-impacts-constant-reporters.pdf"))
```

### Social performance constant reporters vs. all

```{r}
#| fig-width: 15
#| fig-height: 16

library(jsonlite) # TEMP: append to analyses.qmd and requirements.txt

constant_reporters <- jsonlite::fromJSON(
  here::here("data/raw/datasets/constant_reporters.json")
) # TEMP: decide on where JSON file should be stored


selected_indicators <- c(
  "Employee turnover [%]",
  "Employees covered by collective bargaining agreements [%]",
  "# Training hours per employee",
  "# Days lost to work-related injuries, ill health, and fatalities among employees",
  "Female employees in top managment [%]",
  "Gender pay gap [%]",
  "Annual remuneration ratio [%]",
  "# Complaints filed by own workforce",
  "Amount of fines, penalties, and compensation for damages as result of incidents and complaints [USD]"
)

# bins
n_breaks  <- c(0, 150, 300, Inf)
n_labels  <- c("≤150", "≤300", ">300")

# prepare base data + rename labels
temp_renamed <- temp |>
  select(firm, year, topic, id, label_specification, value_final) |>
  mutate(
    label_specification = recode(label_specification, !!!renamer_soc),
    label_specification = factor(label_specification, levels = unname(renamer_soc))
  ) |>
  filter(topic == "soc", label_specification %in% selected_indicators) |>
  drop_na(value_final)

# create mapping from display label -> indicator id
tmp <- temp_renamed |>
  distinct(id, label_specification)

name_to_id_map <- stats::setNames(tmp$id, as.character(tmp$label_specification))

# flag constant reporters observation-wise
temp_flagged <- temp_renamed |>
  rowwise() |>
  mutate(
    ind_id = name_to_id_map[as.character(label_specification)],
    is_constant_reporter = if (!is.na(ind_id) && ind_id %in% names(constant_reporters)) {
      firm %in% constant_reporters[[ind_id]]
    } else {
      FALSE
    }
  ) |>
  ungroup() |>
  select(-ind_id)

# grouping info
grouping_info <- temp_flagged |>
  filter(is_constant_reporter == TRUE) |>
  group_by(label_specification, year) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(
    n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
    n_group = factor(n_group, levels = n_labels)
  )

p10_fun <- function(x, label) {
  if (label != "Gender pay gap [%]") pmax(stats::quantile(x, 0.10, na.rm = TRUE), 1)
  else stats::quantile(x, 0.10, na.rm = TRUE)
}

# summaries for ALL
p_all <- temp_flagged |>
  group_by(label_specification, year) |>
  summarise(
    p10    = p10_fun(value_final, dplyr::first(label_specification)),
    q1     = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3     = quantile(value_final, 0.75, na.rm = TRUE),
    p90    = quantile(value_final, 0.90, na.rm = TRUE),
    group  = "All companies",
    .groups = "drop"
  )

# summaries for CONSTANT REPORTERS only
p_constant <- temp_flagged |>
  filter(is_constant_reporter == TRUE) |>
  group_by(label_specification, year) |>
  mutate(
    lower_bound = quantile(value_final, 0.00, na.rm = TRUE),
    upper_bound = quantile(value_final, 0.99, na.rm = TRUE),
    value_final = pmin(pmax(value_final, lower_bound), upper_bound)
  ) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10    = p10_fun(value_final, dplyr::first(label_specification)),
    q1     = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3     = quantile(value_final, 0.75, na.rm = TRUE),
    p90    = quantile(value_final, 0.90, na.rm = TRUE),
    group  = "Constant reporters",
    .groups = "drop"
  )

p_soc <- bind_rows(p_all, p_constant) |>
  left_join(grouping_info, by = c("label_specification", "year")) |>
  mutate(
    n_group = factor(n_group, levels = n_labels),
    group   = factor(group, levels = c("All companies", "Constant reporters"))
  )

# n label (counts for constant reporters)
n_values <- temp_flagged |>
  filter(is_constant_reporter == TRUE) |>
  group_by(label_specification) |>
  summarise(n = n(), .groups = "drop")

# indicators that should stay on linear scale
linear_indicators <- c(
  "Employee turnover [%]",
  "Employees covered by collective bargaining agreements [%]",
  "Female employees in top managment [%]",
  "Gender pay gap [%]",
  "Annual remuneration ratio [%]"
)

plotting <- function(indicator) {

  data_filtered <- p_soc |>
    filter(label_specification == indicator) |>
    mutate(n_group = factor(n_group, levels = n_labels)) |>
    tidyr::complete(
      year,
      label_specification,
      group = c("All companies", "Constant reporters"),
      n_group = factor(n_labels, levels = n_labels)
    )

  n_label <- n_values |>
    filter(label_specification == indicator) |>
    pull(n)

  label_n <- paste0("italic(n)==", n_label)

  # y max (only used for log-scale)
  y_max <- max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1

  p <- ggplot(data_filtered, aes(x = year)) +
    geom_line(
      data = dplyr::filter(data_filtered, !is.na(median)),
      aes(y = median, color = group, linetype = group),
      linewidth = 1
    )

  if (indicator %in% linear_indicators) {
    p <- p +
      scale_y_continuous(
        limits = c(
          min(data_filtered$p10, na.rm = TRUE) * 0.9,
          max(data_filtered$p90, na.rm = TRUE) * 1.1
        ),
        labels = scales::comma_format()
      )
  } else {
    p <- p +
      scale_y_continuous(
        trans = "log10",
        labels = scales::comma_format(),
        limits = c(1, y_max),
        oob = scales::oob_keep
      )
  }

  p <- p +
    scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
    scale_color_manual(
      values = c("All companies" = "grey50", "Constant reporters" = "#6F2E5D"),
      name = NULL
    ) +
    scale_linetype_manual(
      values = c("All companies" = "dashed", "Constant reporters" = "solid"),
      name = NULL
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
      panel.grid.minor.y = element_blank(),
      axis.ticks = element_line(color = "grey"),
      axis.ticks.length = unit(0.15, "cm"),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16)
    ) +
    labs(
      title = stringr::str_wrap(indicator, width = 45),
      x = "",
      y = parse(text = "")
    )

  label_grob <- grid::textGrob(
    label = parse(text = label_n),
    x = unit(0.98, "npc"),
    y = unit(0.95, "npc"),
    just = c("right", "bottom"),
    gp = grid::gpar(fontsize = 14)
  )

  p <- p + annotation_custom(label_grob)

  ggsave(
    filename = here("fig", paste0("fig-soc-constant-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
    plot = p,
    width = 5, height = 5, dpi = 400, units = "in", device = "png"
  )

  p
}

plots <- lapply(selected_indicators, plotting)

pfinal <- patchwork::wrap_plots(plots) +
  patchwork::plot_layout(ncol = 3, guides = "collect") +
  patchwork::plot_annotation(tag_levels = "a") &
  theme(
    plot.tag = element_text(face = "bold", size = 20),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.justification = "center",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 16)
  )

ggsave(
  here("fig", "fig-soc-impacts-constant-reporters.pdf"),
  pfinal,
  width = 16, height = 17, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-soc-impacts-constant-reporters.pdf"))
```

### Governance performance constant reporters vs. all

```{r}
#| fig-width: 15
#| fig-height: 5.5

library(jsonlite) # TEMP: append to analyses.qmd and requirements.txt

constant_reporters <- jsonlite::fromJSON(
  here::here("data/raw/datasets/constant_reporters.json")
) # TEMP: decide on where JSON file should be stored

selected_indicators <- c(
  "Independent board members [%]",
  "Lobbying expenses [USD]",
  "# Days to pay invoice"
)

# define bins
n_breaks <- c(0, 150, 300, Inf)
n_labels <- c("≤150", "≤300", ">300")

# prepare base data + rename labels
temp_renamed <- temp |>
  select(firm, year, topic, id, label_specification, value_final) |>
  mutate(
    label_specification = recode(label_specification, !!!renamer_gov),
    label_specification = factor(label_specification, levels = unname(renamer_gov))
  ) |>
  filter(topic == "gov", label_specification %in% selected_indicators) |>
  drop_na(value_final)

# create mapping from display label -> indicator id
tmp <- temp_renamed |>
  distinct(id, label_specification)

name_to_id_map <- stats::setNames(tmp$id, as.character(tmp$label_specification))

# flag constant reporters per observation (firm x year x indicator)
temp_flagged <- temp_renamed |>
  rowwise() |>
  mutate(
    ind_id = name_to_id_map[as.character(label_specification)],
    is_constant_reporter = if (!is.na(ind_id) && ind_id %in% names(constant_reporters)) {
      firm %in% constant_reporters[[ind_id]]
    } else {
      FALSE
    }
  ) |>
  ungroup() |>
  select(-ind_id)

# grouping info for constant reporters
grouping_info <- temp_flagged |>
  filter(is_constant_reporter == TRUE) |>
  group_by(label_specification, year) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(
    n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
    n_group = factor(n_group, levels = n_labels)
  )

# summaries: ALL companies
p_all <- temp_flagged |>
  group_by(label_specification, year) |>
  summarise(
    p10 = pmax(quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    group = "All companies",
    .groups = "drop"
  )

# summaries: CONSTANT REPORTERS
p_constant <- temp_flagged |>
  filter(is_constant_reporter == TRUE) |>
  group_by(label_specification, year) |>
  mutate(
    lower_bound = quantile(value_final, 0.00, na.rm = TRUE),
    upper_bound = quantile(value_final, 0.99, na.rm = TRUE),
    value_final = pmin(pmax(value_final, lower_bound), upper_bound)
  ) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10 = pmax(quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    group = "Constant reporters",
    .groups = "drop"
  )

p_gov <- bind_rows(p_all, p_constant) |>
  left_join(grouping_info, by = c("label_specification", "year")) |>
  mutate(
    n_group = factor(n_group, levels = n_labels),
    group = factor(group, levels = c("All companies", "Constant reporters"))
  )

# n label (constant reporters counts)
n_values <- temp_flagged |>
  filter(is_constant_reporter == TRUE) |>
  group_by(label_specification) |>
  summarise(n = n(), .groups = "drop")

# indicators that should stay on linear scale
linear_indicators <- c("Independent board members [%]")

plotting <- function(indicator) {

  data_filtered <- p_gov |>
    filter(label_specification == indicator) |>
    mutate(n_group = factor(n_group, levels = n_labels)) |>
    tidyr::complete(
      year,
      label_specification,
      group = c("All companies", "Constant reporters"),
      n_group = factor(n_labels, levels = n_labels)
    )

  n_label <- n_values |>
    filter(label_specification == indicator) |>
    pull(n)
  
  # if no match, set to 0
  if (length(n_label) == 0 || is.na(n_label)) n_label <- 0

  label_n <- sprintf("italic(n)==%s", n_label)

  y_max <- max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1

  p <- ggplot(data_filtered, aes(x = year)) +
    geom_line(
      data = dplyr::filter(data_filtered, !is.na(median)),
      aes(y = median, color = group, linetype = group),
      linewidth = 1
    )

  if (indicator %in% linear_indicators) {
    p <- p +
      scale_y_continuous(
        limits = c(
          min(data_filtered$p10, na.rm = TRUE) * 0.9,
          max(data_filtered$p90, na.rm = TRUE) * 1.1
        ),
        labels = scales::comma_format()
      )
  } else {
    p <- p +
      scale_y_continuous(
        trans = "log10",
        labels = scales::comma_format(),
        limits = c(1, y_max),
        oob = scales::oob_keep
      )
  }

  p <- p +
    scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
    scale_color_manual(
      values = c("All companies" = "grey50", "Constant reporters" = "#485DA5"),
      name = NULL
    ) +
    scale_linetype_manual(
      values = c("All companies" = "dashed", "Constant reporters" = "solid"),
      name = NULL
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
      panel.grid.minor.y = element_blank(),
      axis.ticks = element_line(color = "grey"),
      axis.ticks.length = unit(0.15, "cm"),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16)
    ) +
    labs(
      title = stringr::str_wrap(indicator, width = 45),
      x = "",
      y = parse(text = "")
    )

  label_grob <- grid::textGrob(
    label = parse(text = label_n),
    x = unit(0.98, "npc"),
    y = unit(0.95, "npc"),
    just = c("right", "bottom"),
    gp = grid::gpar(fontsize = 14)
  )

  p <- p + annotation_custom(label_grob)

  ggsave(
    filename = here("fig", paste0("fig-gov-constant-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
    plot = p,
    width = 5, height = 5, dpi = 400, units = "in", device = "png"
  )

  p
}

plots <- lapply(selected_indicators, plotting)

pfinal <- patchwork::wrap_plots(plots) +
  patchwork::plot_layout(ncol = 3, guides = "collect") +
  patchwork::plot_annotation(tag_levels = "a") &
  theme(
    plot.tag = element_text(face = "bold", size = 20),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.justification = "center",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 16)
  )

ggsave(
  here("fig", "fig-gov-impacts-constant-reporters.pdf"),
  pfinal,
  width = 16, height = 6, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-gov-impacts-constant-reporters.pdf"))

```

### Transparency high-quality vs. all

```{r}
#| fig-width: 10
#| fig-height: 4

library(jsonlite) # TEMP: append to analyes-qmd and requirements.txt

std_levels <- c(
    "Climate change",
    "Pollution",
    "Water",
    "Biodiversity",
    "Circular economy",
    "Own workforce",
    "Consumers",
    "Governance",
    "Business conduct"
)

colors <- c(
    "Climate change"   = "#1C4A29",
    "Pollution"        = "#2E6F40",
    "Water"            = "#419759",
    "Biodiversity"     = "#55C173",
    "Circular economy" = "#6AEC8E",
    "Own workforce"    = "#6F2E5D",
    "Consumers"        = "#D661B5",
    "Governance"       = "#485DA5",
    "Business conduct" = "#6D81CD"
)

df <- df |> mutate(standard_topic = ifelse(standard_topic == "General", "Governance", standard_topic))

# load list of high-quality company-years
hq_json <- jsonlite::fromJSON(
  here::here("data/raw/datasets/top_quality_documents_yearly.json")
) # TEMP: decide on storage location 

hq_keys <- hq_json$top_quality_documents_yearly

# helper: compute "avg # ESG indicators reported" per year and ESRS topic
calc_transparency_bars <- function(data) {
  data |>
    dplyr::mutate(std = factor(standard_topic, levels = std_levels)) |>
    dplyr::group_by(firm, year, std) |>
    dplyr::summarise(count_reported = sum(exists), .groups = "drop") |>
    dplyr::group_by(year, std) |>
    dplyr::summarise(avg_reported = mean(count_reported), .groups = "drop")
}

# ensure company_year exists once
if (!("company_year" %in% names(df))) {
  df <- df |>
    dplyr::mutate(company_year = paste0(firm, "_", year))
}

# Panel a - all companies
p_all <- df |>
  calc_transparency_bars() |>
  ggplot(aes(x = factor(year), y = avg_reported, fill = std)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_discrete(breaks = as.character(seq(2015, 2023, by = 2))) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 220)) +
  labs(x = NULL, y = "# ESG indicators", fill = "ESRS Topic") +
  ggtitle("All companies") +
  theme(
    plot.title = element_text(size = 13),
    text = element_text(size = 11),
    axis.line = element_line(linewidth = 0.5),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    axis.ticks = element_line(),
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 11, face = "bold")
  )

# Panel b - high-quality company-years
p_hq <- df |>
  dplyr::filter(company_year %in% hq_keys) |>
  calc_transparency_bars() |>
  ggplot(aes(x = factor(year), y = avg_reported, fill = std)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_discrete(breaks = as.character(seq(2015, 2023, by = 2))) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 220)) +
  labs(x = NULL, y = NULL, fill = "ESRS Topic") +
  ggtitle("High quality company-years") +
  theme(
    plot.title = element_text(size = 13),
    text = element_text(size = 11),
    axis.text.x = element_text(),
    axis.ticks.x = element_line(),
    axis.line.x = element_line(size = 0.3),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 11, face = "bold")
  )

pfinal_hq_transp <- (p_all | p_hq) +
  patchwork::plot_layout(guides = "collect") +
  patchwork::plot_annotation(tag_levels = "a") &
  theme(
    plot.tag = element_text(face = "bold"),
    legend.position = "right",
    legend.direction = "vertical"
  )

ggsave(
  here("fig", "fig-transparency_trend_high_quality.pdf"),
  pfinal_hq_transp,
  width = 10, height = 4, dpi = 400, units = "in", device = "pdf"
)

knitr::include_graphics(here("fig/fig-transparency_trend_high_quality.pdf"))

```

```{r}
#| fig-width: 5
#| fig-height: 6

library(jsonlite) # TEMP: append to analyses.qmd and requirements.txt

shapes_line <- c(
    "Climate change"   = 16,
    "Pollution"        = 17,
    "Water"            = 22,
    "Biodiversity"     = 4,
    "Circular economy" = 23,
    "Own workforce"    = 16,
    "Consumers"        = 17,
    "Governance"       = 1,
    "Business conduct" = 17
)

shapes_scatter <- c(
    "Climate change"   = 16,
    "Pollution"        = 17,
    "Water"            = 0,
    "Biodiversity"     = 4,
    "Circular economy" = 5,
    "Own workforce"    = 16,
    "Consumers"        = 17,
    "Governance"       = 1,
    "Business conduct" = 17
)

temp <- temp |> mutate(standard_topic = ifelse(standard_topic == "General", "Governance", standard_topic))

# load list of high-quality company-years
hq_json <- jsonlite::fromJSON(
  here::here("data/raw/datasets/top_quality_documents_yearly.json")
) # TEMP: decide on storage location for file
hq_keys <- hq_json$top_quality_documents_yearly

# ensure company_year exists (base-pipe friendly)
if (!("company_year" %in% names(temp))) {
  temp <- temp |>
    dplyr::mutate(company_year = paste0(firm, "_", year))
}

# filter to HQ firm-years
df_hq <- temp |>
  dplyr::filter(company_year %in% hq_keys)

# add aggregate benchmark transparency to time series plot (p2) - HQ ONLY
overall_p2_hq <-
  df_hq |>
  dplyr::group_by(firm, year) |>
  dplyr::summarize(transp = mean(exists), .groups = "drop") |>
  dplyr::group_by(year) |>
  dplyr::summarize(transp = mean(transp), .groups = "drop") |>
  dplyr::mutate(
    label = "Overall",
    group = "Overall"
  )

# rebuild p2 - HQ ONLY
p2_hq <-
  df_hq |>
  dplyr::group_by(firm, year, topic, standard, standard_topic) |>
  dplyr::summarize(transp = mean(exists), .groups = "drop") |>
  dplyr::group_by(year, topic, standard, standard_topic) |>
  dplyr::summarize(transp = mean(transp), .groups = "drop") |>
  dplyr::mutate(
    label = paste0(standard_topic),
    group = ifelse(topic == "env", "Environmental", "Social and Governance")
  ) |>
  dplyr::bind_rows(overall_p2_hq) |>
  dplyr::mutate(
    label = factor(label, levels = c("Overall", std_levels))
  ) |>
  ggplot2::ggplot(ggplot2::aes(
    x = factor(year),
    y = transp,
    color = label,
    shape = label,
    group = label
  )) +
  ggplot2::geom_line() +
  ggplot2::geom_point(size = 2, fill = "white", stroke = 1) +
  ggplot2::scale_y_continuous(labels = scales::percent, limits = c(0, 0.6), expand = c(0, 0)) +
  ggplot2::scale_color_manual(values = c("Overall" = "darkgrey", colors)) +
  ggplot2::scale_shape_manual(values = c("Overall" = 16, shapes_line)) +
  ggplot2::labs(x = NULL, y = "Transparency score", color = "ESRS Topic", shape = "ESRS Topic") +
  ggplot2::theme(
    text = ggplot2::element_text(color = "black", size = 10),
    panel.grid.major = ggplot2::element_blank(),
    panel.border = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.line = ggplot2::element_line(color = "black"),
    axis.ticks = ggplot2::element_line(color = "black"),
    legend.key.height = grid::unit(10, "pt"),
    legend.position = "left",
    legend.box.margin = ggplot2::margin(r = -120),
    axis.title.y = ggplot2::element_text(vjust = -55)
  )

# add aggregate benchmark to sectoral scatter plot (p3) - HQ ONLY
overall_p3_hq <-
  df_hq |>
  dplyr::group_by(firm, year) |>
  dplyr::summarize(transp = mean(exists), .groups = "drop") |>
  add_all_firm_data() |>
  dplyr::group_by(primary_sics_sector) |>
  dplyr::summarize(transp = mean(transp), .groups = "drop") |>
  dplyr::filter(!is.na(primary_sics_sector)) |>
  dplyr::mutate(std = "Overall")

sector_order_hq <- overall_p3_hq |>
  dplyr::arrange(transp) |>
  dplyr::pull(primary_sics_sector)

# rebuild p3 - HQ ONLY
p3_hq <-
  df_hq |>
  dplyr::mutate(std = paste0(standard_topic)) |>
  dplyr::group_by(firm, year, topic, std, standard) |>
  dplyr::summarize(transp = mean(exists), .groups = "drop") |>
  add_all_firm_data() |>
  dplyr::group_by(primary_sics_sector, topic, std, standard) |>
  dplyr::summarize(transp = mean(transp), .groups = "drop") |>
  dplyr::filter(!is.na(primary_sics_sector)) |>
  dplyr::bind_rows(overall_p3_hq) |>
  dplyr::mutate(
    primary_sics_sector = factor(primary_sics_sector, levels = sector_order_hq),
    std = factor(std, levels = c("Overall", std_levels))
  ) |>
  ggplot2::ggplot(ggplot2::aes(
    x = transp,
    y = primary_sics_sector,
    color = std,
    shape = std
  )) +
  ggplot2::geom_point(size = 2, stroke = 1) +
  ggplot2::scale_x_continuous(labels = scales::percent, limits = c(0, 0.6), expand = c(0, 0)) +
  ggplot2::scale_color_manual(values = c("Overall" = "darkgrey", colors)) +
  ggplot2::scale_shape_manual(values = c("Overall" = 16, shapes_scatter)) +
  ggplot2::labs(x = "Transparency score", y = "", shape = "ESRS Topic", color = "ESRS Topic") +
  ggplot2::theme(
    text = ggplot2::element_text(color = "black", size = 10),
    panel.border = ggplot2::element_blank(),
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.line = ggplot2::element_line(color = "black"),
    axis.ticks = ggplot2::element_line(color = "black"),
    legend.position = "none"
  )

# combine + export - HQ ONLY
p3_hq <- p3_hq + ggplot2::guides(color = "none", shape = "none")

pfinal_hq <- p2_hq / p3_hq +
  patchwork::plot_layout(
    heights = c(1, 1.4),
    guides = "collect"
  ) +
  patchwork::plot_annotation(tag_levels = "a") &
  ggplot2::theme(
    plot.tag = ggplot2::element_text(face = "bold"),
    plot.caption = ggtext::element_textbox_simple(hjust = 0),
    plot.caption.position = "plot",
    legend.position = "right",
    legend.justification = "center"
  )

ggsave(
  here("fig", "fig-transparency_high_quality.pdf"),
  pfinal_hq,
  width = 8, height = 6, dpi = 400, units = "in", device = "pdf"
)

knitr::include_graphics(here("fig/fig-transparency_high_quality.pdf"))

```
```{r}
#| fig-width: 9.5
#| fig-height: 5

library(jsonlite) # TEMP: append to analyses.qmd and requirements.txt

# Updated violin plot with top vs bottom 10% (vs middle 80%) - HIGH QUALITY firm-years only
if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {

    # load list of high-quality company-years
    hq_json <- jsonlite::fromJSON(
        here::here("data/raw/datasets/top_quality_documents_yearly.json")
    ) # TEMP: decide on storage location for file
    hq_keys <- hq_json$top_quality_documents_yearly

    # ensure company_year exists in df (base-pipe friendly)
    if (!("company_year" %in% names(df))) {
        df <- df |>
            dplyr::mutate(company_year = paste0(firm, "_", year))
    }

    esg_rating_map <- c(
        "A+" = 12, "A" = 11, "A-" = 10,
        "B+" = 9,  "B" = 8,  "B-" = 7,
        "C+" = 6,  "C" = 5,  "C-" = 4,
        "D+" = 3,  "D" = 2,  "D-" = 1
    )

    # reverse-coded
    controversy_score_map <- c(
        "A+" = 1, "A" = 2, "A-" = 3,
        "B+" = 4, "B" = 5, "B-" = 6,
        "C+" = 7, "C" = 8, "C-" = 9,
        "D+" = 10, "D" = 11, "D-" = 12
    )

    msci_company_rating_map <- c(
        "AAA" = 7, "AA" = 6, "A" = 5,
        "BBB" = 4, "BB" = 3, "B" = 2,
        "CCC" = 1
    )

    # Base summarization and firm data enrichment (HQ only)
    df2_enriched <- df |>
        dplyr::filter(company_year %in% hq_keys) |>
        dplyr::group_by(firm, year) |>
        dplyr::summarize(transp = mean(exists), .groups = "drop") |>
        add_all_firm_data()

    # Continue transformation
    df2 <- df2_enriched |>
        dplyr::select(
            firm, year, primary_sics_sector, transp,
            mkcapUSD_lagged, salesUSD,
            ref_esg_score, ref_esg_controversies_score_lagged,
            msci_company_rating_lagged,
            reprisk_rating_num_lagged
        ) |>
        dplyr::mutate(
            ref_esg_score_n = dplyr::recode(ref_esg_score, !!!esg_rating_map),
            ref_esg_controversies_score_n = dplyr::recode(ref_esg_controversies_score_lagged, !!!controversy_score_map),
            msci_company_rating_n = dplyr::recode(msci_company_rating_lagged, !!!msci_company_rating_map)
        )

    make_grouped_df <- function(df, quant_col) {
        df_grouped <- df |>
            # Compute quantile thresholds yearly
            dplyr::group_by(year) |>
            dplyr::mutate(
                q10 = stats::quantile(.data[[quant_col]], 0.1, na.rm = TRUE),
                q90 = stats::quantile(.data[[quant_col]], 0.9, na.rm = TRUE),
                group = dplyr::case_when(
                    is.na(.data[[quant_col]]) ~ NA_character_,
                    .data[[quant_col]] >= q90 ~ "Top 10%",
                    .data[[quant_col]] <= q10 ~ "Bottom 10%",
                    TRUE ~ "Middle 80%"
                )
            ) |>
            dplyr::ungroup() |>
            dplyr::select(-q10, -q90)

        return(df_grouped)
    }

    df_mkcap_plot <- make_grouped_df(df2, "mkcapUSD_lagged")
    # df_sales_plot <- make_grouped_df(df2, "salesUSD")
    # df_rating_plot <- make_grouped_df(df2, "ref_esg_score_n")
    df_controversies_plot <- make_grouped_df(df2, "ref_esg_controversies_score_n")
    df_msci_plot <- make_grouped_df(df2, "msci_company_rating_n")
    # df_reprisk_rating_plot <- make_grouped_df(df2, "reprisk_rating_num_lagged")

    # Function  to compare means
    compare_means <- function(df, group_var = "group", target_group = "Top 10%", reference_group = "Middle 80%") {
        variable_sym <- rlang::sym(group_var)

        x <- df |>
            dplyr::filter((!!variable_sym) == target_group) |>
            dplyr::pull(transp) |>
            stats::na.omit()
        y <- df |>
            dplyr::filter((!!variable_sym) == reference_group) |>
            dplyr::pull(transp) |>
            stats::na.omit()

        ttest <- stats::t.test(x = x, y = y, alternative = "two.sided")

        stars <- dplyr::case_when(
            ttest$p.value < 0.001 ~ "***",
            ttest$p.value < 0.01 ~ "**",
            ttest$p.value < 0.05 ~ "*",
            TRUE ~ ""
        )

        sprintf("diff = %.1f p.p.%s", (ttest$estimate[1] - ttest$estimate[2]) * 100, stars)
    }

    # Function to plot violins
    make_violin_plot <- function(df, x_label, bracket_y = c(0.65, 0.7)) {
        df <- df |> tidyr::drop_na(transp, group)

        ggplot2::ggplot(df, ggplot2::aes(x = factor(group, levels = c("Bottom 10%", "Middle 80%", "Top 10%")), y = transp)) +
            ggplot2::geom_violin(ggplot2::aes(fill = group), alpha = 0.5) +
            ggplot2::geom_boxplot(width = 0.1, outlier.shape = NA) +
            ggplot2::scale_fill_manual(values = blue_shades) +
            ggplot2::scale_y_continuous(labels = scales::percent, limits = c(0, 0.75), expand = c(0, 0)) +
            ggplot2::labs(x = x_label, y = "Transparency score") +
            ggplot2::theme_minimal(base_size = 10) +
            ggplot2::theme(
                panel.grid.major = ggplot2::element_blank(),
                panel.border = ggplot2::element_blank(),
                axis.line = ggplot2::element_line(linewidth = 0.25),
                axis.ticks = ggplot2::element_line(linewidth = 0.25),
                legend.position = "none",
                axis.text.y = if (x_label == "Market capitalization") ggplot2::element_text() else ggplot2::element_blank(),
                axis.title.y = if (x_label == "Market capitalization") ggplot2::element_text() else ggplot2::element_blank()
            ) +
            ggpubr::geom_bracket(
                xmin = c("Bottom 10%", "Top 10%"),
                xmax = c("Middle 80%", "Middle 80%"),
                y.position = bracket_y,
                label = c(
                    compare_means(df, "group", "Bottom 10%", "Middle 80%"),
                    compare_means(df, "group", "Top 10%", "Middle 80%")
                )
            )
    }

    blue_shades <- c(
        "Bottom 10%" = "#bdd7e7",
        "Middle 80%" = "#6baed6",
        "Top 10%" = "#2171b5"
    )

    violin_mkcap <- make_violin_plot(df_mkcap_plot, "Market capitalization")
    # violin_sales <- make_violin_plot(df_sales_plot, "Revenue")
    # violin_rating <- make_violin_plot(df_rating_plot, "ESG rating")
    violin_controversies <- make_violin_plot(df_controversies_plot, "ESG controversies score")
    violin_msci_rating <- make_violin_plot(df_msci_plot, "ESG rating")
    # violin_reprisk_rating <- make_violin_plot(df_reprisk_rating_plot, "RepRisk Rating")
    
    p <- violin_mkcap + violin_msci_rating + violin_controversies +
        patchwork::plot_layout(ncol = 3) +
        patchwork::plot_annotation(tag_levels = "a") &
        ggplot2::theme(
            plot.tag = ggplot2::element_text(face = "bold"),
            plot.caption = ggtext::element_textbox_simple(hjust = 0),
            plot.caption.position = "plot"
        )

    p
    
    ggplot2::ggsave(
        here::here("fig", "fig-heterogeneity-companies-high-quality.pdf"),
        p,
        width = 9.5, height = 5, dpi = 300, units = "in", device = "pdf"
    )
    
    knitr::include_graphics(here::here("fig/fig-heterogeneity-companies-high-quality.pdf"))
}

```

### Environmental performance high-quality vs. all

```{r}
#| fig-width: 15
#| fig-height: 16

library(jsonlite) # TEMP: append to analyses.qmd and requirements.txt

# load list of high-quality company-years
hq_json <- jsonlite::fromJSON(
  here::here("data/raw/datasets/top_quality_documents_yearly.json")
) # TEMP: decide on location of storage for list

hq_keys <- hq_json$top_quality_documents_yearly

selected_indicators <- c(
  "Total scope 1 emissions [tCO₂eq]",
  "Total scope 2 emissions (market-based) [tCO₂eq]",
  "Total scope 3 emissions [tCO₂eq]",
  "Category 6 Travel [tCO₂eq]",
  "Energy consumption [MWh]",
  "Renewable sources in total energy consumption [%]",
  "Water consumption [m³]",
  "Total waste generated [t]",
  "Non-recycled waste [%]"
)

n_breaks <- c(0, 150, 300, Inf)
n_labels <- c("≤150", "≤300", ">300")

# base data (ALL) + flag high-quality company-years
temp_flagged <- temp |>
  dplyr::select(company_year, firm, year, topic, label_specification, value_final) |>
  dplyr::mutate(
    label_specification = dplyr::recode(label_specification, !!!renamer_env),
    label_specification = factor(label_specification, levels = unname(renamer_env)),
    is_high_quality = company_year %in% hq_keys
  ) |>
  dplyr::filter(topic == "env", label_specification %in% selected_indicators) |>
  tidyr::drop_na(value_final)

# grouping info for HQ only
grouping_info <- temp_flagged |>
  dplyr::filter(is_high_quality == TRUE) |>
  dplyr::group_by(label_specification, year) |>
  dplyr::summarise(n = dplyr::n(), .groups = "drop") |>
  dplyr::mutate(
    n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
    n_group = factor(n_group, levels = n_labels)
  )

# summaries for ALL
p_all <- temp_flagged |>
  dplyr::group_by(label_specification, year) |>
  dplyr::summarise(
    p10 = pmax(stats::quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1  = stats::quantile(value_final, 0.25, na.rm = TRUE),
    median = stats::quantile(value_final, 0.50, na.rm = TRUE),
    q3  = stats::quantile(value_final, 0.75, na.rm = TRUE),
    p90 = stats::quantile(value_final, 0.90, na.rm = TRUE),
    group = "All companies",
    .groups = "drop"
  )

# summaries for HIGH QUALITY company-years
p_hq <- temp_flagged |>
  dplyr::filter(is_high_quality == TRUE) |>
  dplyr::group_by(label_specification, year) |>
  dplyr::mutate(
    lower_bound = stats::quantile(value_final, 0.00, na.rm = TRUE),
    upper_bound = stats::quantile(value_final, 0.99, na.rm = TRUE),
    value_final = pmin(pmax(value_final, lower_bound), upper_bound)
  ) |>
  dplyr::ungroup() |>
  dplyr::group_by(label_specification, year) |>
  dplyr::summarise(
    p10 = pmax(stats::quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1  = stats::quantile(value_final, 0.25, na.rm = TRUE),
    median = stats::quantile(value_final, 0.50, na.rm = TRUE),
    q3  = stats::quantile(value_final, 0.75, na.rm = TRUE),
    p90 = stats::quantile(value_final, 0.90, na.rm = TRUE),
    group = "High quality firm-years",
    .groups = "drop"
  )

p_env <- dplyr::bind_rows(p_all, p_hq) |>
  dplyr::left_join(grouping_info, by = c("label_specification", "year")) |>
  dplyr::mutate(
    n_group = factor(n_group, levels = n_labels),
    group = factor(group, levels = c("All companies", "High quality firm-years"))
  )

# n label (HQ counts)
n_values <- temp_flagged |>
  dplyr::filter(is_high_quality == TRUE) |>
  dplyr::group_by(label_specification) |>
  dplyr::summarise(n = dplyr::n(), .groups = "drop")

# emissions share y-limit
emission_indicators <- c(
  "Total scope 1 emissions [tCO₂eq]",
  "Total scope 2 emissions (market-based) [tCO₂eq]",
  "Total scope 3 emissions [tCO₂eq]",
  "Category 6 Travel [tCO₂eq]"
)

shared_emission_max <- p_env |>
  dplyr::filter(label_specification %in% emission_indicators) |>
  dplyr::summarise(ymax = max(p90, q3, na.rm = TRUE)) |>
  dplyr::pull(ymax) * 1.1

plotting <- function(indicator) {

  indicator_title <- switch(indicator,
    "Total scope 1 emissions [tCO₂eq]" = expression(Total~scope~1~emissions~"[tCO"[2]*"eq]"),
    "Total scope 2 emissions (market-based) [tCO₂eq]" = expression(Total~scope~2~emissions~"[tCO"[2]*"eq]"),
    "Total scope 3 emissions [tCO₂eq]" = expression(Total~Scope~3~emissions~"[tCO"[2]*"eq]"),
    "Category 6 Travel [tCO₂eq]" = expression(Scope~3~emissions~(Travel)~"[tCO"[2]*"eq]"),
    "Energy consumption [MWh]" = expression(Energy~consumption~"[MWh]"),
    "Renewable sources in total energy consumption [%]" = expression(atop("Renewable sources in total","energy consumption [%]")),
    "Water consumption [m³]" = expression(Water~consumption~"[m"^3*"]"),
    "Total waste generated [t]" = expression(Total~waste~generated~"[t]"),
    "Non-recycled waste [%]" = expression("Non-recycled waste [%]"),
    expression(.(indicator))
  )

  data_filtered <- p_env |>
    dplyr::filter(label_specification == indicator) |>
    dplyr::mutate(n_group = factor(n_group, levels = n_labels)) |>
    tidyr::complete(
      year,
      label_specification,
      group = c("All companies", "High quality firm-years"),
      n_group = factor(n_labels, levels = n_labels)
    )

  n_label <- n_values |>
    dplyr::filter(label_specification == indicator) |>
    dplyr::pull(n)
  if (length(n_label) == 0 || is.na(n_label)) n_label <- 0
  label_n <- sprintf("italic(n)==%s", n_label)

  y_max <- if (indicator %in% emission_indicators) {
    shared_emission_max
  } else {
    max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1
  }

  p <- ggplot2::ggplot(data_filtered, ggplot2::aes(x = year))

  if (indicator == "Category 6 Travel [tCO₂eq]") {
    p <- p +
      ggplot2::geom_rect(
        data = tibble::tibble(xmin = 2020, xmax = 2023, ymin = 1, ymax = 12000000),
        ggplot2::aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
        fill = "grey85", alpha = 0.3, inherit.aes = FALSE
      ) +
      ggplot2::annotate(
        "text",
        x = 2020.3, y = 500000,
        label = "COVID-19\npandemic",
        hjust = 0, size = 4, color = "grey30"
      )
  }

  p <- p +
    ggplot2::geom_line(
      data = dplyr::filter(data_filtered, !is.na(median)),
      ggplot2::aes(y = median, color = group, linetype = group),
      linewidth = 1
    )

  if (indicator %in% c("Renewable sources in total energy consumption [%]",
                       "Non-recycled waste [%]")) {
    p <- p +
      ggplot2::scale_y_continuous(
        limits = c(min(data_filtered$p10, na.rm = TRUE) * 0.9,
                   max(data_filtered$p90, na.rm = TRUE) * 1.1),
        labels = scales::comma_format()
      )
  } else {
    p <- p +
      ggplot2::scale_y_continuous(
        trans = "log10",
        labels = scales::comma_format(),
        limits = c(1, y_max),
        oob = scales::oob_keep
      )
  }

  p <- p +
    ggplot2::scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
    ggplot2::scale_color_manual(
      values = c("All companies" = "grey50", "High quality firm-years" = "#2E6F40"),
      name = NULL
    ) +
    ggplot2::scale_linetype_manual(
      values = c("All companies" = "dashed", "High quality firm-years" = "solid"),
      name = NULL
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      plot.title = ggplot2::element_text(size = 14, hjust = 0.5),
      panel.border = ggplot2::element_rect(color = "grey", fill = NA, linewidth = 0.8),
      panel.grid.major.x = ggplot2::element_blank(),
      panel.grid.minor.x = ggplot2::element_blank(),
      panel.grid.major.y = ggplot2::element_line(linetype = "dotted", color = "gray60"),
      panel.grid.minor.y = ggplot2::element_blank(),
      axis.ticks = ggplot2::element_line(color = "grey"),
      axis.ticks.length = grid::unit(0.15, "cm"),
      axis.text.x = ggplot2::element_text(size = 16),
      axis.text.y = ggplot2::element_text(size = 16)
    ) +
    ggplot2::labs(
      title = bquote(.(indicator_title)),
      x = "",
      y = parse(text = "")
    )

  label_grob <- grid::textGrob(
    label = parse(text = label_n),
    x = grid::unit(0.98, "npc"),
    y = grid::unit(0.95, "npc"),
    just = c("right", "bottom"),
    gp = grid::gpar(fontsize = 14)
  )

  p <- p + ggplot2::annotation_custom(label_grob)

  ggplot2::ggsave(
    filename = here::here("fig", paste0("fig-hq-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
    plot = p,
    width = 5, height = 5, dpi = 400, units = "in", device = "png"
  )

  p
}

plots <- lapply(selected_indicators, plotting)

pfinal <- patchwork::wrap_plots(plots) +
  patchwork::plot_layout(ncol = 3, guides = "collect") +
  patchwork::plot_annotation(tag_levels = "a") &
  ggplot2::theme(
    plot.tag = ggplot2::element_text(face = "bold", size = 20),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.justification = "center",
    legend.title = ggplot2::element_text(size = 16, face = "bold"),
    legend.text = ggplot2::element_text(size = 16)
  )

ggplot2::ggsave(
  here::here("fig", "fig-env-impacts-high-quality-firmyears.pdf"),
  pfinal,
  width = 16, height = 15, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here::here("fig/fig-env-impacts-high-quality-firmyears.pdf"))

```

### Social performance high-quality vs. all
```{r}
#| fig-width: 15
#| fig-height: 16

library(jsonlite) # TEMP: append to analyses.qmd and requirements.txt

# load list of high-quality company-years
hq_json <- jsonlite::fromJSON(
  here::here("data/raw/datasets/top_quality_documents_yearly.json")
) # TEMP: decide on location of storage for list

hq_keys <- hq_json$top_quality_documents_yearly

selected_indicators <- c(
  "Employee turnover [%]",
  "Employees covered by collective bargaining agreements [%]",
  "# Training hours per employee",
  "# Days lost to work-related injuries, ill health, and fatalities among employees",
  "Female employees in top managment [%]",
  "Gender pay gap [%]",
  "Annual remuneration ratio [%]",
  "# Complaints filed by own workforce",
  "Amount of fines, penalties, and compensation for damages as result of incidents and complaints [USD]"
)

# bins
n_breaks  <- c(0, 150, 300, Inf)
n_labels  <- c("≤150", "≤300", ">300")

# prepare base data + rename labels
temp_flagged <- temp |>
  select(company_year, firm, year, topic, id, label_specification, value_final) |>
  mutate(
    label_specification = recode(label_specification, !!!renamer_soc),
    label_specification = factor(label_specification, levels = unname(renamer_soc)),
    is_high_quality = company_year %in% hq_keys
  ) |>
  filter(topic == "soc", label_specification %in% selected_indicators) |>
  drop_na(value_final)

# grouping info
grouping_info <- temp_flagged |>
  filter(is_high_quality == TRUE) |>
  group_by(label_specification, year) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(
    n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
    n_group = factor(n_group, levels = n_labels)
  )

p10_fun <- function(x, label) {
  if (label != "Gender pay gap [%]") pmax(stats::quantile(x, 0.10, na.rm = TRUE), 1)
  else stats::quantile(x, 0.10, na.rm = TRUE)
}

# summaries for ALL
p_all <- temp_flagged |>
  group_by(label_specification, year) |>
  summarise(
    p10    = p10_fun(value_final, dplyr::first(label_specification)),
    q1     = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3     = quantile(value_final, 0.75, na.rm = TRUE),
    p90    = quantile(value_final, 0.90, na.rm = TRUE),
    group  = "All companies",
    .groups = "drop"
  )

# summaries for HIGH QUALITY company-years only
p_hq <- temp_flagged |>
  filter(is_high_quality == TRUE) |>
  group_by(label_specification, year) |>
  mutate(
    lower_bound = quantile(value_final, 0.00, na.rm = TRUE),
    upper_bound = quantile(value_final, 0.99, na.rm = TRUE),
    value_final = pmin(pmax(value_final, lower_bound), upper_bound)
  ) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10    = p10_fun(value_final, dplyr::first(label_specification)),
    q1     = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3     = quantile(value_final, 0.75, na.rm = TRUE),
    p90    = quantile(value_final, 0.90, na.rm = TRUE),
    group  = "High quality firm-years",
    .groups = "drop"
  )

p_soc <- bind_rows(p_all, p_hq) |>
  left_join(grouping_info, by = c("label_specification", "year")) |>
  mutate(
    n_group = factor(n_group, levels = n_labels),
    group   = factor(group, levels = c("All companies", "High quality firm-years"))
  )

# n label (counts for high-quality firm-years)
n_values <- temp_flagged |>
  filter(is_high_quality == TRUE) |>
  group_by(label_specification) |>
  summarise(n = n(), .groups = "drop")

# indicators that should stay on linear scale
linear_indicators <- c(
  "Employee turnover [%]",
  "Employees covered by collective bargaining agreements [%]",
  "Female employees in top managment [%]",
  "Gender pay gap [%]",
  "Annual remuneration ratio [%]"
)

plotting <- function(indicator) {

  data_filtered <- p_soc |>
    filter(label_specification == indicator) |>
    mutate(n_group = factor(n_group, levels = n_labels)) |>
    tidyr::complete(
      year,
      label_specification,
      group = c("All companies", "High quality firm-years"),
      n_group = factor(n_labels, levels = n_labels)
    )

  n_label <- n_values |>
    filter(label_specification == indicator) |>
    pull(n)
  if (length(n_label) == 0 || is.na(n_label)) n_label <- 0

  label_n <- paste0("italic(n)==", n_label)

  # y max (only used for log-scale)
  y_max <- max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1

  p <- ggplot(data_filtered, aes(x = year)) +
    geom_line(
      data = dplyr::filter(data_filtered, !is.na(median)),
      aes(y = median, color = group, linetype = group),
      linewidth = 1
    )

  if (indicator %in% linear_indicators) {
    p <- p +
      scale_y_continuous(
        limits = c(
          min(data_filtered$p10, na.rm = TRUE) * 0.9,
          max(data_filtered$p90, na.rm = TRUE) * 1.1
        ),
        labels = scales::comma_format()
      )
  } else {
    p <- p +
      scale_y_continuous(
        trans = "log10",
        labels = scales::comma_format(),
        limits = c(1, y_max),
        oob = scales::oob_keep
      )
  }

  p <- p +
    scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
    scale_color_manual(
      values = c("All companies" = "grey50", "High quality firm-years" = "#6F2E5D"),
      name = NULL
    ) +
    scale_linetype_manual(
      values = c("All companies" = "dashed", "High quality firm-years" = "solid"),
      name = NULL
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
      panel.grid.minor.y = element_blank(),
      axis.ticks = element_line(color = "grey"),
      axis.ticks.length = unit(0.15, "cm"),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16)
    ) +
    labs(
      title = stringr::str_wrap(indicator, width = 45),
      x = "",
      y = parse(text = "")
    )

  label_grob <- grid::textGrob(
    label = parse(text = label_n),
    x = unit(0.98, "npc"),
    y = unit(0.95, "npc"),
    just = c("right", "bottom"),
    gp = grid::gpar(fontsize = 14)
  )

  p <- p + annotation_custom(label_grob)

  ggsave(
    filename = here("fig", paste0("fig-soc-hq-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
    plot = p,
    width = 5, height = 5, dpi = 400, units = "in", device = "png"
  )

  p
}

plots <- lapply(selected_indicators, plotting)

pfinal <- patchwork::wrap_plots(plots) +
  patchwork::plot_layout(ncol = 3, guides = "collect") +
  patchwork::plot_annotation(tag_levels = "a") &
  theme(
    plot.tag = element_text(face = "bold", size = 20),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.justification = "center",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 16)
  )

ggsave(
  here("fig", "fig-soc-impacts-high-quality-firmyears.pdf"),
  pfinal,
  width = 16, height = 17, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-soc-impacts-high-quality-firmyears.pdf"))

```

### Governance performance high-quality vs. all
```{r}
#| fig-width: 15
#| fig-height: 5.5

library(jsonlite) # TEMP: append to analyses.qmd and requirements.txt

# load list of high-quality company-years
hq_json <- jsonlite::fromJSON(
  here::here("data/raw/datasets/top_quality_documents_yearly.json")
) # TEMP: decide on location of storage for list

hq_keys <- hq_json$top_quality_documents_yearly

selected_indicators <- c(
  "Independent board members [%]",
  "Lobbying expenses [USD]",
  "# Days to pay invoice"
)

# define bins
n_breaks <- c(0, 150, 300, Inf)
n_labels <- c("≤150", "≤300", ">300")

# prepare base data + rename labels
temp_flagged <- temp |>
  select(company_year, firm, year, topic, id, label_specification, value_final) |>
  mutate(
    label_specification = recode(label_specification, !!!renamer_gov),
    label_specification = factor(label_specification, levels = unname(renamer_gov)),
    is_high_quality = company_year %in% hq_keys
  ) |>
  filter(topic == "gov", label_specification %in% selected_indicators) |>
  drop_na(value_final)

# grouping info for high-quality company-years
grouping_info <- temp_flagged |>
  filter(is_high_quality == TRUE) |>
  group_by(label_specification, year) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(
    n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
    n_group = factor(n_group, levels = n_labels)
  )

# summaries: ALL companies
p_all <- temp_flagged |>
  group_by(label_specification, year) |>
  summarise(
    p10 = pmax(quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    group = "All companies",
    .groups = "drop"
  )

# summaries: HIGH QUALITY company-years
p_hq <- temp_flagged |>
  filter(is_high_quality == TRUE) |>
  group_by(label_specification, year) |>
  mutate(
    lower_bound = quantile(value_final, 0.00, na.rm = TRUE),
    upper_bound = quantile(value_final, 0.99, na.rm = TRUE),
    value_final = pmin(pmax(value_final, lower_bound), upper_bound)
  ) |>
  ungroup() |>
  group_by(label_specification, year) |>
  summarise(
    p10 = pmax(quantile(value_final, 0.10, na.rm = TRUE), 1),
    q1 = quantile(value_final, 0.25, na.rm = TRUE),
    median = quantile(value_final, 0.50, na.rm = TRUE),
    q3 = quantile(value_final, 0.75, na.rm = TRUE),
    p90 = quantile(value_final, 0.90, na.rm = TRUE),
    group = "High quality firm-years",
    .groups = "drop"
  )

p_gov <- bind_rows(p_all, p_hq) |>
  left_join(grouping_info, by = c("label_specification", "year")) |>
  mutate(
    n_group = factor(n_group, levels = n_labels),
    group = factor(group, levels = c("All companies", "High quality firm-years"))
  )

# n label (high-quality firm-years counts)
n_values <- temp_flagged |>
  filter(is_high_quality == TRUE) |>
  group_by(label_specification) |>
  summarise(n = n(), .groups = "drop")

# indicators that should stay on linear scale
linear_indicators <- c("Independent board members [%]")

plotting <- function(indicator) {

  data_filtered <- p_gov |>
    filter(label_specification == indicator) |>
    mutate(n_group = factor(n_group, levels = n_labels)) |>
    tidyr::complete(
      year,
      label_specification,
      group = c("All companies", "High quality firm-years"),
      n_group = factor(n_labels, levels = n_labels)
    )

  n_label <- n_values |>
    filter(label_specification == indicator) |>
    pull(n)

  # if no match, set to 0
  if (length(n_label) == 0 || is.na(n_label)) n_label <- 0

  label_n <- sprintf("italic(n)==%s", n_label)

  y_max <- max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1

  p <- ggplot(data_filtered, aes(x = year)) +
    geom_line(
      data = dplyr::filter(data_filtered, !is.na(median)),
      aes(y = median, color = group, linetype = group),
      linewidth = 1
    )

  if (indicator %in% linear_indicators) {
    p <- p +
      scale_y_continuous(
        limits = c(
          min(data_filtered$p10, na.rm = TRUE) * 0.9,
          max(data_filtered$p90, na.rm = TRUE) * 1.1
        ),
        labels = scales::comma_format()
      )
  } else {
    p <- p +
      scale_y_continuous(
        trans = "log10",
        labels = scales::comma_format(),
        limits = c(1, y_max),
        oob = scales::oob_keep
      )
  }

  p <- p +
    scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
    scale_color_manual(
      values = c("All companies" = "grey50", "High quality firm-years" = "#485DA5"),
      name = NULL
    ) +
    scale_linetype_manual(
      values = c("All companies" = "dashed", "High quality firm-years" = "solid"),
      name = NULL
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
      panel.grid.minor.y = element_blank(),
      axis.ticks = element_line(color = "grey"),
      axis.ticks.length = unit(0.15, "cm"),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16)
    ) +
    labs(
      title = stringr::str_wrap(indicator, width = 45),
      x = "",
      y = parse(text = "")
    )

  label_grob <- grid::textGrob(
    label = parse(text = label_n),
    x = unit(0.98, "npc"),
    y = unit(0.95, "npc"),
    just = c("right", "bottom"),
    gp = grid::gpar(fontsize = 14)
  )

  p <- p + annotation_custom(label_grob)

  ggsave(
    filename = here("fig", paste0("fig-gov-hq-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
    plot = p,
    width = 5, height = 5, dpi = 400, units = "in", device = "png"
  )

  p
}

plots <- lapply(selected_indicators, plotting)

pfinal <- patchwork::wrap_plots(plots) +
  patchwork::plot_layout(ncol = 3, guides = "collect") +
  patchwork::plot_annotation(tag_levels = "a") &
  theme(
    plot.tag = element_text(face = "bold", size = 20),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.justification = "center",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 16)
  )

ggsave(
  here("fig", "fig-gov-impacts-high-quality-firmyears.pdf"),
  pfinal,
  width = 16, height = 6, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-gov-impacts-high-quality-firmyears.pdf"))

```

### Transparency based on disclosure-detections

```{r}
#| fig-width: 6
#| fig-height: 4

std_levels <- c(
    "Climate change",
    "Pollution",
    "Water",
    "Biodiversity",
    "Circular economy",
    "Own workforce",
    "Consumers",
    "Governance",
    "Business conduct"
)

colors <- c(
    "Climate change"   = "#1C4A29",
    "Pollution"        = "#2E6F40",
    "Water"            = "#419759",
    "Biodiversity"     = "#55C173",
    "Circular economy" = "#6AEC8E",
    "Own workforce"    = "#6F2E5D",
    "Consumers"        = "#D661B5",
    "Governance"       = "#485DA5",
    "Business conduct" = "#6D81CD"
)

df <- df |> mutate(standard_topic = ifelse(standard_topic == "General", "Governance", standard_topic))

# create detection flag: TRUE if model output has non-null "value", 0 otherwise
df <- df |>
  dplyr::mutate(
    detection = !is.na(model_output) &
      !stringr::str_detect(model_output, "'value':\\s*None(?!')")
  )

# helper: compute avg # ESG indicators detected per year and ESRS topic
calc_transparency_bars <- function(data) {
  data |>
    dplyr::mutate(std = factor(standard_topic, levels = std_levels)) |>
    dplyr::group_by(firm, year, std) |>
    dplyr::summarise(count_detected = sum(detection), .groups = "drop") |>
    dplyr::group_by(year, std) |>
    dplyr::summarise(avg_detected = mean(count_detected), .groups = "drop")
}

p_all <- df |>
  calc_transparency_bars() |>
  ggplot(aes(x = factor(year), y = avg_detected, fill = std)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_discrete(breaks = as.character(seq(2015, 2023, by = 2))) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 220)) +
  labs(x = NULL, y = "# ESG indicators", fill = "ESRS Topic") +
  theme(
    text = element_text(size = 11),
    axis.line = element_line(linewidth = 0.5),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    axis.ticks = element_line(),
    legend.position = "right",
    legend.direction = "vertical",
    strip.background = element_blank(),
    strip.text = element_text(size = 11, face = "bold")
  )

ggsave(
  here("fig", "fig-transparency_trend_detection.pdf"),
  p_all,
  width = 6, height = 4, dpi = 400, units = "in", device = "pdf"
)

knitr::include_graphics(here("fig/fig-transparency_trend_detection.pdf"))

```

```{r}
#| fig-width: 5
#| fig-height: 6

shapes_line <- c(
    "Climate change"   = 16,
    "Pollution"        = 17,
    "Water"            = 22,
    "Biodiversity"     = 4,
    "Circular economy" = 23,
    "Own workforce"    = 16,
    "Consumers"        = 17,
    "Governance"       = 1,
    "Business conduct" = 17
)

shapes_scatter <- c(
    "Climate change"   = 16,
    "Pollution"        = 17,
    "Water"            = 0,
    "Biodiversity"     = 4,
    "Circular economy" = 5,
    "Own workforce"    = 16,
    "Consumers"        = 17,
    "Governance"       = 1,
    "Business conduct" = 17
)

# aggregate benchmark transparency for time series plot (p2)
overall_p2 <-
  df |>
  dplyr::group_by(firm, year) |>
  dplyr::summarize(transp = mean(detection), .groups = "drop") |>
  dplyr::group_by(year) |>
  dplyr::summarize(transp = mean(transp), .groups = "drop") |>
  dplyr::mutate(
    label = "Overall",
    group = "Overall"
  )

# p2: transparency time series by ESRS topic
p2 <-
  df |>
  dplyr::group_by(firm, year, topic, standard, standard_topic) |>
  dplyr::summarize(transp = mean(detection), .groups = "drop") |>
  dplyr::group_by(year, topic, standard, standard_topic) |>
  dplyr::summarize(transp = mean(transp), .groups = "drop") |>
  dplyr::mutate(
    label = paste0(standard_topic),
    group = ifelse(topic == "env", "Environmental", "Social and Governance")
  ) |>
  dplyr::bind_rows(overall_p2) |>
  dplyr::mutate(
    label = factor(label, levels = c("Overall", std_levels))
  ) |>
  ggplot2::ggplot(ggplot2::aes(
    x = factor(year),
    y = transp,
    color = label,
    shape = label,
    group = label
  )) +
  ggplot2::geom_line() +
  ggplot2::geom_point(size = 2, fill = "white", stroke = 1) +
  ggplot2::scale_y_continuous(labels = scales::percent, limits = c(0, 0.6), expand = c(0, 0)) +
  ggplot2::scale_color_manual(values = c("Overall" = "darkgrey", colors)) +
  ggplot2::scale_shape_manual(values = c("Overall" = 16, shapes_line)) +
  ggplot2::labs(x = NULL, y = "Transparency score", color = "ESRS Topic", shape = "ESRS Topic") +
  ggplot2::theme(
    text = ggplot2::element_text(color = "black", size = 10),
    panel.grid.major = ggplot2::element_blank(),
    panel.border = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.line = ggplot2::element_line(color = "black"),
    axis.ticks = ggplot2::element_line(color = "black"),
    legend.key.height = grid::unit(10, "pt"),
    legend.position = "left",
    legend.box.margin = ggplot2::margin(r = -120),
    axis.title.y = ggplot2::element_text(vjust = -55)
  )

# aggregate benchmark for sectoral scatter plot (p3)
overall_p3 <-
  df |>
  dplyr::group_by(firm, year) |>
  dplyr::summarize(transp = mean(detection), .groups = "drop") |>
  add_all_firm_data() |>
  dplyr::group_by(primary_sics_sector) |>
  dplyr::summarize(transp = mean(transp), .groups = "drop") |>
  dplyr::filter(!is.na(primary_sics_sector)) |>
  dplyr::mutate(std = "Overall")

sector_order <- overall_p3 |>
  dplyr::arrange(transp) |>
  dplyr::pull(primary_sics_sector)

# p3: sectoral scatter
p3 <-
  df |>
  dplyr::mutate(std = paste0(standard_topic)) |>
  dplyr::group_by(firm, year, topic, std, standard) |>
  dplyr::summarize(transp = mean(detection), .groups = "drop") |>
  add_all_firm_data() |>
  dplyr::group_by(primary_sics_sector, topic, std, standard) |>
  dplyr::summarize(transp = mean(transp), .groups = "drop") |>
  dplyr::filter(!is.na(primary_sics_sector)) |>
  dplyr::bind_rows(overall_p3) |>
  dplyr::mutate(
    primary_sics_sector = factor(primary_sics_sector, levels = sector_order),
    std = factor(std, levels = c("Overall", std_levels))
  ) |>
  ggplot2::ggplot(ggplot2::aes(
    x = transp,
    y = primary_sics_sector,
    color = std,
    shape = std
  )) +
  ggplot2::geom_point(size = 2, stroke = 1) +
  ggplot2::scale_x_continuous(labels = scales::percent, limits = c(0, 0.6), expand = c(0, 0)) +
  ggplot2::scale_color_manual(values = c("Overall" = "darkgrey", colors)) +
  ggplot2::scale_shape_manual(values = c("Overall" = 16, shapes_scatter)) +
  ggplot2::labs(x = "Transparency score", y = "", shape = "ESRS Topic", color = "ESRS Topic") +
  ggplot2::theme(
    text = ggplot2::element_text(color = "black", size = 10),
    panel.border = ggplot2::element_blank(),
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.line = ggplot2::element_line(color = "black"),
    axis.ticks = ggplot2::element_line(color = "black"),
    legend.position = "none"
  )

# combine + export
p3 <- p3 + ggplot2::guides(color = "none", shape = "none")

pfinal <- p2 / p3 +
  patchwork::plot_layout(
    heights = c(1, 1.4),
    guides = "collect"
  ) +
  patchwork::plot_annotation(tag_levels = "a") &
  ggplot2::theme(
    plot.tag = ggplot2::element_text(face = "bold"),
    plot.caption = ggtext::element_textbox_simple(hjust = 0),
    plot.caption.position = "plot",
    legend.position = "right",
    legend.justification = "center"
  )

ggsave(
  here("fig", "fig-transparency_detection.pdf"),
  pfinal,
  width = 8, height = 6, dpi = 400, units = "in", device = "pdf"
)

knitr::include_graphics(here("fig/fig-transparency_detection.pdf"))

```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Transparency trends conditional on market capitalization"`

```{r}
#| fig-width: 10
#| fig-height: 4

std_levels <- c(
  "Climate change", "Pollution", "Water", "Biodiversity", "Circular economy",
  "Own workforce", "Consumers", "Governance", "Business conduct"
)

colors <- c(
  "Climate change"   = "#1C4A29",
  "Pollution"        = "#2E6F40",
  "Water"            = "#419759",
  "Biodiversity"     = "#55C173",
  "Circular economy" = "#6AEC8E",
  "Own workforce"    = "#6F2E5D",
  "Consumers"        = "#D661B5",
  "Governance"       = "#485DA5",
  "Business conduct" = "#6D81CD"
)

temp <- temp |>
  mutate(standard_topic = ifelse(standard_topic == "General", "Governance", standard_topic))

# create market-cap terciles within each year
mkcap_groups_yearly <- temp |>
  distinct(firm, year, mkcapUSD_lagged) |>
  group_by(year) |>
  filter(!is.na(mkcapUSD_lagged)) |>
  mutate(
    q33 = quantile(mkcapUSD_lagged, probs = 1/3, na.rm = TRUE),
    q67 = quantile(mkcapUSD_lagged, probs = 2/3, na.rm = TRUE),
    mkcap_group = case_when(
      mkcapUSD_lagged <= q33 ~ "Market-cap tercile: bottom",
      mkcapUSD_lagged >= q67 ~ "Market-cap tercile: top",
      TRUE ~ "Market-cap tercile: middle"
    )
  ) |>
  ungroup() |>
  select(firm, year, mkcap_group) |>
  distinct()

# join group back to main data
temp <- temp |>
  left_join(mkcap_groups_yearly, by = c("firm", "year"))

mkcap_levels <- c(
  "Market-cap tercile: bottom",
  "Market-cap tercile: middle",
  "Market-cap tercile: top"
)

# --- build three panels (same logic as your rating panels) ---
p_mkcap_small <- temp |>
  filter(mkcap_group == mkcap_levels[1]) |>
  mutate(std = factor(standard_topic, levels = std_levels)) |>
  group_by(firm, year, std) |>
  summarize(count_reported = sum(exists), .groups = "drop") |>
  group_by(year, std) |>
  summarize(avg_reported = mean(count_reported), .groups = "drop") |>
  ggplot(aes(x = factor(year), y = avg_reported, fill = std)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_discrete(breaks = as.character(seq(2015, 2023, by = 2))) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 220)) +
  labs(x = NULL, y = "# ESG indicators", fill = "ESRS Topic") +
  ggtitle("Market-cap tercile: bottom") +
  theme(
    plot.title = element_text(size = 13),
    text = element_text(size = 11),
    axis.line = element_line(linewidth = 0.5),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    axis.ticks = element_line(),
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 11, face = "bold")
  )

p_mkcap_mid <- temp |>
  filter(mkcap_group == mkcap_levels[2]) |>
  mutate(std = factor(standard_topic, levels = std_levels)) |>
  group_by(firm, year, std) |>
  summarize(count_reported = sum(exists), .groups = "drop") |>
  group_by(year, std) |>
  summarize(avg_reported = mean(count_reported), .groups = "drop") |>
  ggplot(aes(x = factor(year), y = avg_reported, fill = std)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_discrete(breaks = as.character(seq(2015, 2023, by = 2))) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 220)) +
  labs(x = NULL, y = NULL, fill = "ESRS Topic") +
  ggtitle("Market-cap tercile: middle") +
  theme(
    plot.title = element_text(size = 13),
    text = element_text(size = 11),
    axis.text.x = element_text(),
    axis.ticks.x = element_line(),
    axis.line.x = element_line(linewidth = 0.3),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 11, face = "bold")
  )

p_mkcap_large <- temp |>
  filter(mkcap_group == mkcap_levels[3]) |>
  mutate(std = factor(standard_topic, levels = std_levels)) |>
  group_by(firm, year, std) |>
  summarize(count_reported = sum(exists), .groups = "drop") |>
  group_by(year, std) |>
  summarize(avg_reported = mean(count_reported), .groups = "drop") |>
  ggplot(aes(x = factor(year), y = avg_reported, fill = std)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_discrete(breaks = as.character(seq(2015, 2023, by = 2))) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 220)) +
  labs(x = NULL, y = NULL, fill = "ESRS Topic") +
  ggtitle("Market-cap tercile: top") +
  theme(
    plot.title = element_text(size = 13),
    text = element_text(size = 11),
    axis.text.x = element_text(),
    axis.ticks.x = element_line(),
    axis.line.x = element_line(linewidth = 0.3),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 11, face = "bold")
  )

pfinal_mkcap <- (p_mkcap_small | p_mkcap_mid | p_mkcap_large) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    plot.tag = element_text(face = "bold"),
    legend.position = "right",
    legend.direction = "vertical"
  )

ggsave(
  here::here("fig", "fig-transparency_mkcap_terciles.pdf"),
  pfinal_mkcap,
  width = 10, height = 4, dpi = 400, units = "in", device = "pdf"
)

knitr::include_graphics(here::here("fig/fig-transparency_mkcap_terciles.pdf"))

```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Environmental performance top vs bottom 10% - Market capitalization"`

```{r performance-mkcap}
# ESG performance top vs bottom 10% - Market capitalization
#| fig-width: 16
#| fig-height: 15
if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {
  selected_indicators <- c("Total scope 1 emissions [tCO₂eq]",
                          "Total scope 2 emissions (market-based) [tCO₂eq]",
                          "Total scope 3 emissions [tCO₂eq]",
                          #"Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]",
                          "Category 6 Travel [tCO₂eq]",
                          #"Scope 3 emissions (Commuting) [tCO₂eq]",
                          "Energy consumption [MWh]",
                          
                          "Renewable sources in total energy consumption [%]",
                          #"Fossil sources in total energy consumption [%]",
                          "Water consumption [m³]",
                          "Total waste generated [t]",
                          "Non-recycled waste [%]"
                          )

  # Base summarization and firm data enrichment
  df2_enriched <- temp |>
      add_all_firm_data()

  # Finalize transformation
  temp2 <- df2_enriched |>
    mutate(label_specification = recode(label_specification, !!!renamer_env),
          label_specification = factor(label_specification, levels = unname(renamer_env))
          ) |>
    filter(label_specification %in% selected_indicators) |>
    select(firm, year, label_specification, value_final, mkcapUSD_lagged) |> # use lagged mkcap
    rename(mkcapUSD = mkcapUSD_lagged)

  # Assign groups based on market cap percentiles (yearly)
  temp2 <- temp2 |>
    group_by(year) |> 
    mutate(
      top10 = mkcapUSD >= quantile(mkcapUSD, 0.90, na.rm = TRUE),
      bottom10 = mkcapUSD <= quantile(mkcapUSD, 0.10, na.rm = TRUE)
    ) |>
    ungroup() |>
    mutate(
      mkcap_group = case_when(
        top10 ~ "Top 10%",
        bottom10 ~ "Bottom 10%",
        TRUE ~ "Middle 80%"
      ),
      mkcap_group = factor(mkcap_group, levels = c("Top 10%", "Bottom 10%")) # set only for top and bottom 10%
    )

  # define bins 
  n_breaks <- c(0, 150, 300, Inf)
  n_labels <- c("≤150", "≤300", ">300")

  grouping_info <- temp2 |>
    group_by(label_specification, year, mkcap_group) |>
    summarise(
      n = n(),
      .groups = "drop"
    ) |>
    mutate(
      n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
      n_group = factor(n_group, levels = n_labels)
    )

  n_values <- temp2 |>
    filter(!is.na(value_final)) |> # use only non-null observations (shown in plot)
    group_by(label_specification) |> # for all companies regardless of marketcap 
    summarise(n = n(), .groups = "drop")

  # summarise values
  p_summary <- temp2 |>
    group_by(label_specification, year, mkcap_group) |> 
    summarise(
      p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Overall median
  overall_median <- temp2 |> 
    group_by(label_specification, year) |> 
    summarise(
      median = quantile(value_final, 0.50, na.rm = TRUE),
      .groups = "drop"
    ) |> 
    mutate(mkcap_group = "All companies")

  p_summary_all <- bind_rows(p_summary, overall_median) # bind rows

  # set y axis limits independently for emission metrics
  emission_indicators <- c(
    "Total scope 1 emissions [tCO₂eq]",
    "Total scope 2 emissions (market-based) [tCO₂eq]",
    "Total scope 3 emissions [tCO₂eq]",
    "Category 6 Travel [tCO₂eq]"
  )

  shared_emission_max <- p_summary |>
    filter(label_specification %in% emission_indicators) |>
    summarise(ymax = max(p90, q3, na.rm = TRUE)) |>
    pull(ymax) * 1.1


  plotting <- function(indicator) {
    indicator_title <- switch(indicator,
      "Total scope 1 emissions [tCO₂eq]" = expression(Total~scope~1~emissions~"[tCO"[2]*"eq]"),
      "Total scope 2 emissions (market-based) [tCO₂eq]" = expression(Total~scope~2~emissions~"[tCO"[2]*"eq]"),
      "Total scope 3 emissions [tCO₂eq]" = expression(Total~Scope~3~emissions~"[tCO"[2]*"eq]"),
      "Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]" = expression(Scope~3~emissions~(Cloud~computing~and~data~center~services)~"[tCO"[2]*"eq"),
      "Category 6 Travel [tCO₂eq]" = expression(atop(Scope~3~emissions, (Category~6~Travel)~"[tCO"[2]*"eq]")),
      "Scope 3 emissions (Commuting) [tCO₂eq]" = expression(Scope~3~emissions~(Commuting)~"[tCO"[2]*"eq"),
      "Energy consumption [MWh]" = expression(Energy~consumption~"[MWh]"),
      "Renewable sources in total energy consumption [%]" = expression(atop("Renewable sources in total", "energy consumption [%]")),
      "Fossil sources in total energy consumption [%]" = expression(atop("Fossil sources in total", "energy consumption [%]")),
      "Water consumption [m³]" = expression(Water~consumption~"[m"^3*"]"),
      "Total waste generated [t]" = expression(Total~waste~generated~"[t]"),
      "Non-recycled waste [%]" = expression("Non-recycled~waste [%]"),
      expression(.(indicator))
    )
    
  data_filtered <- p_summary_all |> # use p_summary_all
    filter(label_specification == indicator) |>
    left_join(grouping_info, by = c("label_specification", "year", "mkcap_group"))

    n_label <- n_values |>
      filter(label_specification == indicator) |>
      pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

    y_max <- if (indicator %in% emission_indicators) {
      shared_emission_max
    } else {
      max(data_filtered$q3, data_filtered$median, na.rm = TRUE) * 1.1
    }
    
    p <- ggplot(data_filtered, aes(x = year))

    if (indicator == "Scope 3 emissions (Travel) [tCO₂eq]") {
        # add shaded area in the background
        p <- p +
          geom_rect(
            data = tibble(xmin = 2020, xmax = 2023, ymin = 1, ymax = 12000000),
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey85", alpha = 0.3, inherit.aes = FALSE
          ) +

          # add annotation text in the foreground
          annotate(
            "text",
            x = 2020.3, y = 500000,
            label = "COVID-19\npandemic",
            hjust = 0, size = 4, color = "grey30"
          )
    }

    p <- p + 
    geom_line(
        data = data_filtered |> filter(mkcap_group == "All companies"),
        aes(y = median, color = mkcap_group, linetype = mkcap_group),
        linewidth = 1
      ) +
      geom_line(
        data = data_filtered |> filter(mkcap_group %in% c("Top 10%", "Bottom 10%")), # show median for separate groups
        aes(y = median, color = mkcap_group, linetype = mkcap_group),
        size = 1
      ) 
    
      if (indicator %in% c(
      "Renewable sources in total energy consumption [%]",
      "Non-recycled waste [%]"
      )) {
        p <- p +
          scale_y_continuous(
            limits = c(
              min(data_filtered$p10, na.rm = TRUE) * 0.9,
              max(data_filtered$median, na.rm = TRUE) * 1.1
              ),
            labels = scales::comma_format()
            )
        } else {
          p <- p +
            scale_y_continuous(
              trans = "log10",
              labels = scales::comma_format(),
              limits = c(1, y_max),
              oob = scales::oob_keep
            )
        }
    
    p <- p +
      scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
      scale_color_manual(
        values = c("All companies" = "grey50", "Top 10%" = "#2E6F40", "Bottom 10%" = "#C94C4C"), # show only for top / bottom 10%
        name = NULL
      ) +
      scale_linetype_manual(
        values = c("All companies" = "dashed", "Top 10%" = "solid", "Bottom 10%" = "solid"), # show only for top / bottom 10%
        name = NULL
      ) +
    
      theme_minimal() +
      theme(
        plot.title = element_text(size = 14, hjust = 0.5),
        panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
        panel.grid.minor.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        axis.ticks.length = unit(0.15, "cm"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
      ) +
      labs(
        title = bquote(.(indicator_title)),
        x = "",
        y = parse(text = "")
      )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
        )

      p <- p + annotation_custom(label_grob)

      ggsave(
        filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = 'png'
      )

      return(p)
  }


  plots <- lapply(selected_indicators, plotting)

  pfinal <- wrap_plots(plots) +
    plot_layout(
      ncol = 3,
      guides = "collect"
    ) +
    plot_annotation(
      tag_levels = "a"
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 20),
      legend.position = "bottom",
      legend.box = "vertical",
      legend.justification = "center",
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 16)
    )

  ggsave(
      here("fig", "fig-env-impacts-mkcap.pdf"),
      pfinal,
      width = 16, height = 15, dpi = 400, units = "in", device = cairo_pdf
  )

  knitr::include_graphics(here("fig/fig-env-impacts-mkcap.pdf"))
}
```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Social performance top vs bottom 10% - Market capitalization"`

```{r}
#| fig-width: 16
#| fig-height: 15
if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {
  selected_indicators <- c("Employee turnover [%]",
                          "Employees covered by collective bargaining agreements [%]",
                          "# Training hours per employee",
                          "# Days lost to work-related injuries, ill health, and fatalities among employees",
                          #"Employees entitled to take family-related leaves [%]",
                          "Female employees in top managment [%]",
                          "Gender pay gap [%]",
                          "Annual remuneration ratio [%]",
                          "# Complaints filed by own workforce",
                          "Amount of fines, penalties, and compensation for damages as result of incidents and complaints [USD]"
                          #"# Severe human rights incidents connected to own workforce"
                          )

  temp2 <- temp |>
    mutate(label_specification = recode(label_specification, !!!renamer_soc),
          label_specification = factor(label_specification, levels = unname(renamer_soc))
          ) |>
    filter(label_specification %in% selected_indicators) |>
    select(firm, year, label_specification, value_final, mkcapUSD_lagged) |> 
    rename(mkcapUSD = mkcapUSD_lagged)

  # Assign groups based on market cap percentiles (yearly)
  temp2 <- temp2 |>
    group_by(year) |> 
    mutate(
      top10 = mkcapUSD >= quantile(mkcapUSD, 0.90, na.rm = TRUE),
      bottom10 = mkcapUSD <= quantile(mkcapUSD, 0.10, na.rm = TRUE)
    ) |>
    ungroup() |>
    mutate(
      mkcap_group = case_when(
        top10 ~ "Top 10%",
        bottom10 ~ "Bottom 10%",
        TRUE ~ "Middle 80%"
      ),
      mkcap_group = factor(mkcap_group, levels = c("Top 10%", "Bottom 10%"))
    )

  # define bins 
  n_breaks <- c(0, 150, 300, Inf)
  n_labels <- c("≤150", "≤300", ">300")

  grouping_info <- temp2 |>
    group_by(label_specification, year, mkcap_group) |>
    summarise(
      n = n(),
      .groups = "drop"
    ) |>
    mutate(
      n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
      n_group = factor(n_group, levels = n_labels)
    )

  n_values <- temp2 |>
    filter(!is.na(value_final)) |> 
    group_by(label_specification) |>
    summarise(n = n(), .groups = "drop")

  # summarise values
  p_summary <- temp2 |>
    group_by(label_specification, year, mkcap_group) |> 
    summarise(
      p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Overall median
  overall_median <- temp2 |> 
    group_by(label_specification, year) |> 
    summarise(
      median = quantile(value_final, 0.50, na.rm = TRUE),
      .groups = "drop"
    ) |> 
    mutate(mkcap_group = "All companies")

  p_summary_all <- bind_rows(p_summary, overall_median) # bind rows

  plotting <- function(indicator) {

    data_filtered <- p_summary_all |>
    filter(label_specification == indicator) |>
    left_join(grouping_info, by = c("label_specification", "year", "mkcap_group"))

    n_label <- n_values |>
      filter(label_specification == indicator) |>
      pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

    y_max <- if (indicator %in% emission_indicators) {
      shared_emission_max
    } else {
      max(data_filtered$q3, data_filtered$median, na.rm = TRUE) * 1.1
    }
    
    p <- ggplot(data_filtered, aes(x = year))

    p <- p +
      geom_line(
        data = data_filtered |> filter(mkcap_group == "All companies"),
        aes(y = median, color = mkcap_group, linetype = mkcap_group),
        linewidth = 1
      ) +
      geom_line(
        data = data_filtered |> filter(mkcap_group %in% c("Top 10%", "Bottom 10%")), # show median for separate groups
        aes(y = median, color = mkcap_group, linetype = mkcap_group),
        size = 1
      ) 
    
      if (indicator %in% c("Employee turnover [%]",
                          "Employees covered by collective bargaining agreements [%]",
                          "Female employees in top managment [%]",
                          "Gender pay gap [%]",
                          "Annual remuneration ratio [%]")) {
        p <- p + scale_y_continuous(
          limits = c(min(data_filtered$p10, na.rm = TRUE) * 0.9,
                    max(data_filtered$median, na.rm = TRUE) * 1.1),
          labels = scales::comma_format()
          )
      } else {
          p <- p + scale_y_continuous(
            trans = "log10",
            labels = scales::comma_format(),
            limits = c(1, y_max),
            oob = scales::oob_keep
            )
      }
    
      
    p <- p +
      scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
      scale_color_manual(
        values = c("All companies" = "grey50", "Top 10%" = "#2E6F40", "Bottom 10%" = "#C94C4C"),
        name = NULL
      ) +
      scale_linetype_manual(
        values = c("All companies" = "dashed", "Top 10%" = "solid", "Bottom 10%" = "solid"),
        name = NULL
      ) +
      # scale_shape_manual(
      #   values = setNames(c(21, 21, 21), n_labels),
      #   name = "Reporting intensity (n)",
      #   guide = guide_legend(
      #     override.aes = list(
      #       fill = c("white", "#2E6F40", "#2E6F40"),
      #       size = c(3.5, 5.5, 5.5)
      #       )
      #     ),
      #   drop = FALSE
      #   ) +
      # scale_fill_manual(
      #   values = setNames(c("white", "#2E6F40", "#2E6F40"), n_labels),
      #   drop = FALSE
      # ) +
      # scale_size_manual(
      #   values = setNames(c(3.5, 3.5, 5.5), n_labels),
      #   drop = FALSE
      # ) +
      # guides(
      #   fill = "none",
      #   size = "none"
      # ) +

      theme_minimal() +
      theme(
        plot.title = element_text(size = 14, hjust = 0.5),
        panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
        panel.grid.minor.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        axis.ticks.length = unit(0.15, "cm"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
      ) +
      labs(
        title = stringr::str_wrap(indicator, width = 45),
        x = "",
        y = parse(text = "")
      )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
        )

      p <- p + annotation_custom(label_grob)

      ggsave(
        filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = 'png'
      )

      return(p)
  }


  plots <- lapply(selected_indicators, plotting)

  pfinal <- wrap_plots(plots) +
    plot_layout(
      ncol = 3,
      guides = "collect"
    ) +
    plot_annotation(
      tag_levels = "a"
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 20),
      legend.position = "bottom",
      legend.box = "vertical",
      legend.justification = "center",
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 16)
    )

  ggsave(
      here("fig", "fig-soc-impacts-mkcap.pdf"),
      pfinal,
      width = 15, height = 15.5, dpi = 400, units = "in", device = cairo_pdf
  )

  knitr::include_graphics(here("fig/fig-soc-impacts-mkcap.pdf"))
}
```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Governance performance top vs bottom 10% - Market capitalization"`

```{r}
#| fig-width: 15
#| fig-height: 5.5

if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {
  selected_indicators <- c("Independent board members [%]",
                          #"# Confirmed incidents of corruption and bribery",
                          "Lobbying expenses [USD]",
                          #"Political contributions made [EUR]",
                          #"Variable remuneration dependent on sustainability-targets [%]"
                          "# Days to pay invoice"
                          )

  temp2 <- temp |>
    mutate(label_specification = recode(label_specification, !!!renamer_gov),
          label_specification = factor(label_specification, levels = unname(renamer_gov))
          ) |>
    filter(label_specification %in% selected_indicators) |>
    select(firm, year, label_specification, value_final, mkcapUSD_lagged) |> 
    rename(mkcapUSD = mkcapUSD_lagged)

  # Assign groups based on market cap percentiles (yearly)
  temp2 <- temp2 |>
    group_by(year) |> 
    mutate(
      top10 = mkcapUSD >= quantile(mkcapUSD, 0.90, na.rm = TRUE),
      bottom10 = mkcapUSD <= quantile(mkcapUSD, 0.10, na.rm = TRUE)
    ) |>
    ungroup() |>
    mutate(
      mkcap_group = case_when(
        top10 ~ "Top 10%",
        bottom10 ~ "Bottom 10%",
        TRUE ~ "Middle 80%"
      ),
      mkcap_group = factor(mkcap_group, levels = c("Top 10%", "Bottom 10%"))
    )

  # define bins 
  n_breaks <- c(0, 150, 300, Inf)
  n_labels <- c("≤150", "≤300", ">300")

  grouping_info <- temp2 |>
    group_by(label_specification, year, mkcap_group) |>
    summarise(
      n = n(),
      .groups = "drop"
    ) |>
    mutate(
      n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
      n_group = factor(n_group, levels = n_labels)
    )

  n_values <- temp2 |>
    filter(!is.na(value_final)) |> 
    group_by(label_specification) |>
    summarise(n = n(), .groups = "drop")

  # summarise values
  p_summary <- temp2 |>
    group_by(label_specification, year, mkcap_group) |> 
    summarise(
      p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Overall median
  overall_median <- temp2 |> 
    group_by(label_specification, year) |> 
    summarise(
      median = quantile(value_final, 0.50, na.rm = TRUE),
      .groups = "drop"
    ) |> 
    mutate(mkcap_group = "All companies")

  p_summary_all <- bind_rows(p_summary, overall_median) # bind rows

  plotting <- function(indicator) {

    data_filtered <- p_summary_all |>
    filter(label_specification == indicator) |>
    left_join(grouping_info, by = c("label_specification", "year", "mkcap_group"))

    n_label <- n_values |>
      filter(label_specification == indicator) |>
      pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

    y_max <- if (indicator %in% emission_indicators) {
      shared_emission_max
    } else {
      max(data_filtered$q3, data_filtered$median, na.rm = TRUE) * 1.1
    }
    
    p <- ggplot(data_filtered, aes(x = year)) +
      geom_line(
        data = data_filtered |> filter(mkcap_group == "All companies"),
        aes(y = median, color = mkcap_group, linetype = mkcap_group),
        linewidth = 1
      ) +
      geom_line(
        data = data_filtered |> filter(mkcap_group %in% c("Top 10%", "Bottom 10%")), # show median for separate groups
        aes(y = median, color = mkcap_group, linetype = mkcap_group),
        size = 1
      ) 
    
      if (indicator %in% c("Independent board members [%]")) {
        p <- p + scale_y_continuous(
          limits = c(min(data_filtered$p10, na.rm = TRUE) * 0.9,
                    max(data_filtered$median, na.rm = TRUE) * 1.1),
          labels = scales::comma_format()
          )
      } else {
          p <- p + scale_y_continuous(
            trans = "log10",
            labels = scales::comma_format(),
            limits = c(1, y_max),
            oob = scales::oob_keep
            )
      }
    
    p <- p +
      scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
      scale_color_manual(
        values = c("All companies" = "grey50", "Top 10%" = "#2E6F40", "Bottom 10%" = "#C94C4C"),
        name = NULL
      ) +
      scale_linetype_manual(
        values = c("All companies" = "dashed", "Top 10%" = "solid", "Bottom 10%" = "solid"),
        name = NULL
      ) +
      # scale_shape_manual(
      #   values = setNames(c(21, 21, 21), n_labels),
      #   name = "Reporting intensity (n)",
      #   guide = guide_legend(
      #     override.aes = list(
      #       fill = c("white", "#2E6F40", "#2E6F40"),
      #       size = c(3.5, 5.5, 5.5)
      #       )
      #     ),
      #   drop = FALSE
      #   ) +
      # scale_fill_manual(
      #   values = setNames(c("white", "#2E6F40", "#2E6F40"), n_labels),
      #   drop = FALSE
      # ) +
      # scale_size_manual(
      #   values = setNames(c(3.5, 3.5, 5.5), n_labels),
      #   drop = FALSE
      # ) +
      # guides(
      #   fill = "none",
      #   size = "none"
      # ) +

      theme_minimal() +
      theme(
        plot.title = element_text(size = 14, hjust = 0.5),
        panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
        panel.grid.minor.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        axis.ticks.length = unit(0.15, "cm"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
      ) +
      labs(
        title = stringr::str_wrap(indicator, width = 45),
        x = "",
        y = parse(text = "")
      )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
        )

      p <- p + annotation_custom(label_grob)

      ggsave(
        filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = 'png'
      )

      return(p)
  }


  plots <- lapply(selected_indicators, plotting)

  pfinal <- wrap_plots(plots) +
    plot_layout(
      ncol = 3,
      guides = "collect"
    ) +
    plot_annotation(
      tag_levels = "a"
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 20),
      legend.position = "bottom",
      legend.box = "vertical",
      legend.justification = "center",
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 16)
    )

  ggsave(
      here("fig", "fig-gov-impacts-mkcap.pdf"),
      pfinal,
      width = 15, height = 5.5, dpi = 400, units = "in", device = cairo_pdf
  )

  knitr::include_graphics(here("fig/fig-gov-impacts-mkcap.pdf"))
}
```


```{r}
msci_company_rating_map = c(
    "AAA" = 7, "AA" = 6, "A" = 5,
    "BBB" = 4, "BB" = 3, "B" = 2,
    "CCC" = 1
  )
```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Environmental performance top vs bottom 10% - ESG rating"`

```{r performance-rating}
# ESG performance top vs bottom 10% - ESG rating
if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {
  #| fig-width: 16
  #| fig-height: 15


  selected_indicators <- c("Total scope 1 emissions [tCO₂eq]",
                          "Total scope 2 emissions (market-based) [tCO₂eq]",
                          "Total scope 3 emissions [tCO₂eq]",
                          #"Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]",
                          "Category 6 Travel [tCO₂eq]",
                          #"Scope 3 emissions (Commuting) [tCO₂eq]",
                          "Energy consumption [MWh]",
                          
                          "Renewable sources in total energy consumption [%]",
                          #"Fossil sources in total energy consumption [%]",
                          "Water consumption [m³]",
                          "Total waste generated [t]",
                          "Non-recycled waste [%]"
                          )

  temp2 <- temp |>
    mutate(label_specification = recode(label_specification, !!!renamer_env),
          label_specification = factor(label_specification, levels = unname(renamer_env))
          )
  
   temp2 <- temp2 |>
    mutate(
      msci_company_rating_n = recode(msci_company_rating_lagged, !!!msci_company_rating_map)
    ) |> 
    filter(label_specification %in% selected_indicators) |> 
    select(firm, year, label_specification, value_final, msci_company_rating_n) # use lagged MSCI rating

  # Assign groups based on rating percentiles (yearly)
  temp2 <- temp2 |>
    group_by(year) |> 
    mutate(
      top10 = msci_company_rating_n >= quantile(msci_company_rating_n, 0.90, na.rm = TRUE),
      bottom10 = msci_company_rating_n <= quantile(msci_company_rating_n, 0.10, na.rm = TRUE)
    ) |>
    ungroup() |>
    mutate(
      rating_group = case_when(
        top10 ~ "Top 10%",
        bottom10 ~ "Bottom 10%",
        TRUE ~ "Middle 80%"
      ),
      rating_group = factor(rating_group, levels = c("Top 10%", "Bottom 10%")) # set only for top and bottom 10%
    )

  # define bins 
  n_breaks <- c(0, 150, 300, Inf)
  n_labels <- c("≤150", "≤300", ">300")

  grouping_info <- temp2 |>
    group_by(label_specification, year, rating_group) |>
    summarise(
      n = n(),
      .groups = "drop"
    ) |>
    mutate(
      n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
      n_group = factor(n_group, levels = n_labels)
    )

  n_values <- temp2 |>
    filter(!is.na(value_final)) |> # use only non-null observations (shown in plot)
    group_by(label_specification) |> # for all companies regardless of rating 
    summarise(n = n(), .groups = "drop")

  # summarise values
  p_summary <- temp2 |>
    group_by(label_specification, year, rating_group) |> 
    summarise(
      p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Overall median
  overall_median <- temp2 |> 
    group_by(label_specification, year) |> 
    summarise(
      median = quantile(value_final, 0.50, na.rm = TRUE),
      .groups = "drop"
    ) |> 
    mutate(rating_group = "All companies")

  p_summary_all <- bind_rows(p_summary, overall_median) # bind rows

  # set y axis limits independently for emission metrics
  emission_indicators <- c(
    "Total scope 1 emissions [tCO₂eq]",
    "Total scope 2 emissions (market-based) [tCO₂eq]",
    "Total scope 3 emissions [tCO₂eq]",
    "Category 6 Travel [tCO₂eq]"
  )

  shared_emission_max <- p_summary |>
    filter(label_specification %in% emission_indicators) |>
    summarise(ymax = max(p90, q3, na.rm = TRUE)) |>
    pull(ymax) * 1.1


  plotting <- function(indicator) {
    indicator_title <- switch(indicator,
      "Total scope 1 emissions [tCO₂eq]" = expression(Total~scope~1~emissions~"[tCO"[2]*"eq]"),
      "Total scope 2 emissions (market-based) [tCO₂eq]" = expression(Total~scope~2~emissions~"[tCO"[2]*"eq]"),
      "Total scope 3 emissions [tCO₂eq]" = expression(Total~Scope~3~emissions~"[tCO"[2]*"eq]"),
      "Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]" = expression(Scope~3~emissions~(Cloud~computing~and~data~center~services)~"[tCO"[2]*"eq"),
      "Category 6 Travel [tCO₂eq]" = expression(atop(Scope~3~emissions, (Category~6~Travel)~"[tCO"[2]*"eq]")),
      "Scope 3 emissions (Commuting) [tCO₂eq]" = expression(Scope~3~emissions~(Commuting)~"[tCO"[2]*"eq"),
      "Energy consumption [MWh]" = expression(Energy~consumption~"[MWh]"),
      "Renewable sources in total energy consumption [%]" = expression(atop("Renewable sources in total", "energy consumption [%]")),
      "Fossil sources in total energy consumption [%]" = expression(atop("Fossil sources in total", "energy consumption [%]")),
      "Water consumption [m³]" = expression(Water~consumption~"[m"^3*"]"),
      "Total waste generated [t]" = expression(Total~waste~generated~"[t]"),
      "Non-recycled waste [%]" = expression("Non-recycled~waste [%]"),
      expression(.(indicator))
    )
    
  data_filtered <- p_summary_all |> # use p_summary_all
    filter(label_specification == indicator) |>
    left_join(grouping_info, by = c("label_specification", "year", "rating_group"))

    n_label <- n_values |>
      filter(label_specification == indicator) |>
      pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

    y_max <- if (indicator %in% emission_indicators) {
      shared_emission_max
    } else {
      max(data_filtered$q3, data_filtered$median, na.rm = TRUE) * 1.1
    }
    
    p <- ggplot(data_filtered, aes(x = year))

    if (indicator == "Scope 3 emissions (Travel) [tCO₂eq]") {
        # add shaded area in the background
        p <- p +
          geom_rect(
            data = tibble(xmin = 2020, xmax = 2023, ymin = 1, ymax = 12000000),
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey85", alpha = 0.3, inherit.aes = FALSE
          ) +

          # add annotation text in the foreground
          annotate(
            "text",
            x = 2020.3, y = 500000,
            label = "COVID-19\npandemic",
            hjust = 0, size = 4, color = "grey30"
          )
    }

    p <- p + 
    geom_line(
        data = data_filtered |> filter(rating_group == "All companies"),
        aes(y = median, color = rating_group, linetype = rating_group),
        linewidth = 1
      ) +
      geom_line(
        data = data_filtered |> filter(rating_group %in% c("Top 10%", "Bottom 10%")), # show median for separate groups
        aes(y = median, color = rating_group, linetype = rating_group),
        size = 1
      ) 
    
      if (indicator %in% c(
      "Renewable sources in total energy consumption [%]",
      "Non-recycled waste [%]"
      )) {
        p <- p +
          scale_y_continuous(
            limits = c(
              min(data_filtered$p10, na.rm = TRUE) * 0.9,
              max(data_filtered$median, na.rm = TRUE) * 1.1
              ),
            labels = scales::comma_format()
            )
        } else {
          p <- p +
            scale_y_continuous(
              trans = "log10",
              labels = scales::comma_format(),
              limits = c(1, y_max),
              oob = scales::oob_keep
            )
        }
    
    p <- p +
      scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
      scale_color_manual(
        values = c("All companies" = "grey50", "Top 10%" = "#2E6F40", "Bottom 10%" = "#C94C4C"), # show only for top / bottom 10%
        name = NULL
      ) +
      scale_linetype_manual(
        values = c("All companies" = "dashed", "Top 10%" = "solid", "Bottom 10%" = "solid"), # show only for top / bottom 10%
        name = NULL
      ) +
    
      theme_minimal() +
      theme(
        plot.title = element_text(size = 14, hjust = 0.5),
        panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
        panel.grid.minor.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        axis.ticks.length = unit(0.15, "cm"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
      ) +
      labs(
        title = bquote(.(indicator_title)),
        x = "",
        y = parse(text = "")
      )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
        )

      p <- p + annotation_custom(label_grob)

      ggsave(
        filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = 'png'
      )

      return(p)
  }


  plots <- lapply(selected_indicators, plotting)

  pfinal <- wrap_plots(plots) +
    plot_layout(
      ncol = 3,
      guides = "collect"
    ) +
    plot_annotation(
      tag_levels = "a"
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 20),
      legend.position = "bottom",
      legend.box = "vertical",
      legend.justification = "center",
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 16)
    )

  ggsave(
      here("fig", "fig-env-impacts-rating.pdf"),
      pfinal,
      width = 16, height = 15, dpi = 400, units = "in", device = cairo_pdf
  )

  knitr::include_graphics(here("fig/fig-env-impacts-rating.pdf"))
}
```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Social performance top vs bottom 10% - ESG rating"`

```{r}
if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {
  #| fig-width: 16
  #| fig-height: 15

  selected_indicators <- c("Employee turnover [%]",
                          "Employees covered by collective bargaining agreements [%]",
                          "# Training hours per employee",
                          "# Days lost to work-related injuries, ill health, and fatalities among employees",
                          #"Employees entitled to take family-related leaves [%]",
                          "Female employees in top managment [%]",
                          "Gender pay gap [%]",
                          "Annual remuneration ratio [%]",
                          "# Complaints filed by own workforce",
                          "Amount of fines, penalties, and compensation for damages as result of incidents and complaints [USD]"
                          #"# Severe human rights incidents connected to own workforce"
                          )

  temp2 <- temp |>
    mutate(label_specification = recode(label_specification, !!!renamer_soc),
          label_specification = factor(label_specification, levels = unname(renamer_soc))
          ) |>
    mutate(
      msci_company_rating_n = recode(msci_company_rating_lagged, !!!msci_company_rating_map)
    ) |> 
    filter(label_specification %in% selected_indicators) |>
    select(firm, year, label_specification, value_final, msci_company_rating_n)

  # Assign groups based on rating percentiles (yearly)
  temp2 <- temp2 |>
    group_by(year) |> 
    mutate(
      top10 = msci_company_rating_n >= quantile(msci_company_rating_n, 0.90, na.rm = TRUE),
      bottom10 = msci_company_rating_n <= quantile(msci_company_rating_n, 0.10, na.rm = TRUE)
    ) |>
    ungroup() |>
    mutate(
      rating_group = case_when(
        top10 ~ "Top 10%",
        bottom10 ~ "Bottom 10%",
        TRUE ~ "Middle 80%"
      ),
      rating_group = factor(rating_group, levels = c("Top 10%", "Bottom 10%"))
    )

  # define bins 
  n_breaks <- c(0, 150, 300, Inf)
  n_labels <- c("≤150", "≤300", ">300")

  grouping_info <- temp2 |>
    group_by(label_specification, year, rating_group) |>
    summarise(
      n = n(),
      .groups = "drop"
    ) |>
    mutate(
      n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
      n_group = factor(n_group, levels = n_labels)
    )

  n_values <- temp2 |>
    filter(!is.na(value_final)) |> 
    group_by(label_specification) |>
    summarise(n = n(), .groups = "drop")

  # summarise values
  p_summary <- temp2 |>
    group_by(label_specification, year, rating_group) |> 
    summarise(
      p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Overall median
  overall_median <- temp2 |> 
    group_by(label_specification, year) |> 
    summarise(
      median = quantile(value_final, 0.50, na.rm = TRUE),
      .groups = "drop"
    ) |> 
    mutate(rating_group = "All companies")

  p_summary_all <- bind_rows(p_summary, overall_median) # bind rows

  plotting <- function(indicator) {

    data_filtered <- p_summary_all |>
    filter(label_specification == indicator) |>
    left_join(grouping_info, by = c("label_specification", "year", "rating_group"))

    n_label <- n_values |>
      filter(label_specification == indicator) |>
      pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

    y_max <- if (indicator %in% emission_indicators) {
      shared_emission_max
    } else {
      max(data_filtered$q3, data_filtered$median, na.rm = TRUE) * 1.1
    }
    
    p <- ggplot(data_filtered, aes(x = year))

    p <- p +
      geom_line(
        data = data_filtered |> filter(rating_group == "All companies"),
        aes(y = median, color = rating_group, linetype = rating_group),
        linewidth = 1
      ) +
      geom_line(
        data = data_filtered |> filter(rating_group %in% c("Top 10%", "Bottom 10%")), # show median for separate groups
        aes(y = median, color = rating_group, linetype = rating_group),
        size = 1
      ) 
    
      if (indicator %in% c("Employee turnover [%]",
                          "Employees covered by collective bargaining agreements [%]",
                          "Female employees in top managment [%]",
                          "Gender pay gap [%]",
                          "Annual remuneration ratio [%]")) {
        p <- p + scale_y_continuous(
          limits = c(min(data_filtered$p10, na.rm = TRUE) * 0.9,
                    max(data_filtered$median, na.rm = TRUE) * 1.1),
          labels = scales::comma_format()
          )
      } else {
          p <- p + scale_y_continuous(
            trans = "log10",
            labels = scales::comma_format(),
            limits = c(1, y_max),
            oob = scales::oob_keep
            )
      }
    
      
    p <- p +
      scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
      scale_color_manual(
        values = c("All companies" = "grey50", "Top 10%" = "#2E6F40", "Bottom 10%" = "#C94C4C"),
        name = NULL
      ) +
      scale_linetype_manual(
        values = c("All companies" = "dashed", "Top 10%" = "solid", "Bottom 10%" = "solid"),
        name = NULL
      ) +
      # scale_shape_manual(
      #   values = setNames(c(21, 21, 21), n_labels),
      #   name = "Reporting intensity (n)",
      #   guide = guide_legend(
      #     override.aes = list(
      #       fill = c("white", "#2E6F40", "#2E6F40"),
      #       size = c(3.5, 5.5, 5.5)
      #       )
      #     ),
      #   drop = FALSE
      #   ) +
      # scale_fill_manual(
      #   values = setNames(c("white", "#2E6F40", "#2E6F40"), n_labels),
      #   drop = FALSE
      # ) +
      # scale_size_manual(
      #   values = setNames(c(3.5, 3.5, 5.5), n_labels),
      #   drop = FALSE
      # ) +
      # guides(
      #   fill = "none",
      #   size = "none"
      # ) +

      theme_minimal() +
      theme(
        plot.title = element_text(size = 14, hjust = 0.5),
        panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
        panel.grid.minor.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        axis.ticks.length = unit(0.15, "cm"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
      ) +
      labs(
        title = stringr::str_wrap(indicator, width = 45),
        x = "",
        y = parse(text = "")
      )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
        )

      p <- p + annotation_custom(label_grob)

      ggsave(
        filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = 'png'
      )

      return(p)
  }


  plots <- lapply(selected_indicators, plotting)

  pfinal <- wrap_plots(plots) +
    plot_layout(
      ncol = 3,
      guides = "collect"
    ) +
    plot_annotation(
      tag_levels = "a"
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 20),
      legend.position = "bottom",
      legend.box = "vertical",
      legend.justification = "center",
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 16)
    )

  ggsave(
      here("fig", "fig-soc-impacts-rating.pdf"),
      pfinal,
      width = 15, height = 15.5, dpi = 400, units = "in", device = cairo_pdf
  )

  knitr::include_graphics(here("fig/fig-soc-impacts-rating.pdf"))
}
```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Governance performance top vs bottom 10% - ESG rating"`

```{r}
if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {
  #| fig-width: 15
  #| fig-height: 5.5


  selected_indicators <- c("Independent board members [%]",
                          #"# Confirmed incidents of corruption and bribery",
                          "Lobbying expenses [USD]",
                          #"Political contributions made [EUR]",
                          #"Variable remuneration dependent on sustainability-targets [%]"
                          "# Days to pay invoice"
                          )

  temp2 <- temp |>
    mutate(label_specification = recode(label_specification, !!!renamer_gov),
          label_specification = factor(label_specification, levels = unname(renamer_gov))
          ) |>
    mutate(
      msci_company_rating_n = recode(msci_company_rating_lagged, !!!msci_company_rating_map)
    ) |>
    filter(label_specification %in% selected_indicators) |>
    select(firm, year, label_specification, value_final, msci_company_rating_n)

  # Assign groups based on rating percentiles (yearly)
  temp2 <- temp2 |>
    group_by(year) |> 
    mutate(
      top10 = msci_company_rating_n >= quantile(msci_company_rating_n, 0.90, na.rm = TRUE),
      bottom10 = msci_company_rating_n <= quantile(msci_company_rating_n, 0.10, na.rm = TRUE)
    ) |>
    ungroup() |>
    mutate(
      rating_group = case_when(
        top10 ~ "Top 10%",
        bottom10 ~ "Bottom 10%",
        TRUE ~ "Middle 80%"
      ),
      rating_group = factor(rating_group, levels = c("Top 10%", "Bottom 10%"))
    )

  # define bins 
  n_breaks <- c(0, 150, 300, Inf)
  n_labels <- c("≤150", "≤300", ">300")

  grouping_info <- temp2 |>
    group_by(label_specification, year, rating_group) |>
    summarise(
      n = n(),
      .groups = "drop"
    ) |>
    mutate(
      n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
      n_group = factor(n_group, levels = n_labels)
    )

  n_values <- temp2 |>
    filter(!is.na(value_final)) |> 
    group_by(label_specification) |>
    summarise(n = n(), .groups = "drop")

  # summarise values
  p_summary <- temp2 |>
    group_by(label_specification, year, rating_group) |> 
    summarise(
      p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Overall median
  overall_median <- temp2 |> 
    group_by(label_specification, year) |> 
    summarise(
      median = quantile(value_final, 0.50, na.rm = TRUE),
      .groups = "drop"
    ) |> 
    mutate(rating_group = "All companies")

  p_summary_all <- bind_rows(p_summary, overall_median) # bind rows

  plotting <- function(indicator) {

    data_filtered <- p_summary_all |>
    filter(label_specification == indicator) |>
    left_join(grouping_info, by = c("label_specification", "year", "rating_group"))

    n_label <- n_values |>
      filter(label_specification == indicator) |>
      pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

    y_max <- if (indicator %in% emission_indicators) {
      shared_emission_max
    } else {
      max(data_filtered$q3, data_filtered$median, na.rm = TRUE) * 1.1
    }
    
    p <- ggplot(data_filtered, aes(x = year)) +
      geom_line(
        data = data_filtered |> filter(rating_group == "All companies"),
        aes(y = median, color = rating_group, linetype = rating_group),
        linewidth = 1
      ) +
      geom_line(
        data = data_filtered |> filter(rating_group %in% c("Top 10%", "Bottom 10%")), # show median for separate groups
        aes(y = median, color = rating_group, linetype = rating_group),
        size = 1
      ) 
    
      if (indicator %in% c("Independent board members [%]")) {
        p <- p + scale_y_continuous(
          limits = c(min(data_filtered$p10, na.rm = TRUE) * 0.9,
                    max(data_filtered$median, na.rm = TRUE) * 1.1),
          labels = scales::comma_format()
          )
      } else {
          p <- p + scale_y_continuous(
            trans = "log10",
            labels = scales::comma_format(),
            limits = c(1, y_max),
            oob = scales::oob_keep
            )
      }
    
    p <- p +
      scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
      scale_color_manual(
        values = c("All companies" = "grey50", "Top 10%" = "#2E6F40", "Bottom 10%" = "#C94C4C"),
        name = NULL
      ) +
      scale_linetype_manual(
        values = c("All companies" = "dashed", "Top 10%" = "solid", "Bottom 10%" = "solid"),
        name = NULL
      ) +
      # scale_shape_manual(
      #   values = setNames(c(21, 21, 21), n_labels),
      #   name = "Reporting intensity (n)",
      #   guide = guide_legend(
      #     override.aes = list(
      #       fill = c("white", "#2E6F40", "#2E6F40"),
      #       size = c(3.5, 5.5, 5.5)
      #       )
      #     ),
      #   drop = FALSE
      #   ) +
      # scale_fill_manual(
      #   values = setNames(c("white", "#2E6F40", "#2E6F40"), n_labels),
      #   drop = FALSE
      # ) +
      # scale_size_manual(
      #   values = setNames(c(3.5, 3.5, 5.5), n_labels),
      #   drop = FALSE
      # ) +
      # guides(
      #   fill = "none",
      #   size = "none"
      # ) +

      theme_minimal() +
      theme(
        plot.title = element_text(size = 14, hjust = 0.5),
        panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
        panel.grid.minor.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        axis.ticks.length = unit(0.15, "cm"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
      ) +
      labs(
        title = stringr::str_wrap(indicator, width = 45),
        x = "",
        y = parse(text = "")
      )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
        )

      p <- p + annotation_custom(label_grob)

      ggsave(
        filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = 'png'
      )

      return(p)
  }


  plots <- lapply(selected_indicators, plotting)

  pfinal <- wrap_plots(plots) +
    plot_layout(
      ncol = 3,
      guides = "collect"
    ) +
    plot_annotation(
      tag_levels = "a"
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 20),
      legend.position = "bottom",
      legend.box = "vertical",
      legend.justification = "center",
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 16)
    )

  ggsave(
      here("fig", "fig-gov-impacts-rating.pdf"),
      pfinal,
      width = 15, height = 5.5, dpi = 400, units = "in", device = cairo_pdf
  )

  knitr::include_graphics(here("fig/fig-gov-impacts-rating.pdf"))
}
```



```{r}
# reverse-coded
controversy_score_map <- c(
  "A+" = 1, "A" = 2, "A-" = 3,
  "B+" = 4, "B" = 5, "B-" = 6,
  "C+" = 7, "C" = 8, "C-" = 9,
  "D+" = 10, "D" = 11, "D-" = 12
)
```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Environmental performance top vs bottom 10% - ESG controversies score"`
```{r}
# ESG performance top vs bottom 10% - ESG controversies score
if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {
  #| fig-width: 16
  #| fig-height: 15


  selected_indicators <- c("Total scope 1 emissions [tCO₂eq]",
                          "Total scope 2 emissions (market-based) [tCO₂eq]",
                          "Total scope 3 emissions [tCO₂eq]",
                          #"Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]",
                          "Category 6 Travel [tCO₂eq]",
                          #"Scope 3 emissions (Commuting) [tCO₂eq]",
                          "Energy consumption [MWh]",
                          
                          "Renewable sources in total energy consumption [%]",
                          #"Fossil sources in total energy consumption [%]",
                          "Water consumption [m³]",
                          "Total waste generated [t]",
                          "Non-recycled waste [%]"
                          )

  temp2 <- temp |>
    mutate(label_specification = recode(label_specification, !!!renamer_env),
          label_specification = factor(label_specification, levels = unname(renamer_env))
          ) |>
    mutate(
      ref_esg_controversies_score_n = recode(ref_esg_controversies_score_lagged, !!!controversy_score_map)
    ) |> 
    filter(label_specification %in% selected_indicators) |> 
    select(firm, year, label_specification, value_final, ref_esg_controversies_score_n) # use lagged controversies score

  # Assign groups based on controversy score percentiles (yearly)
  temp2 <- temp2 |>
    group_by(year) |> 
    mutate(
      top10 = ref_esg_controversies_score_n >= quantile(ref_esg_controversies_score_n, 0.90, na.rm = TRUE),
      bottom10 = ref_esg_controversies_score_n <= quantile(ref_esg_controversies_score_n, 0.10, na.rm = TRUE)
    ) |>
    ungroup() |>
    mutate(
      controversy_group = case_when(
        top10 ~ "Top 10%",
        bottom10 ~ "Bottom 10%",
        TRUE ~ "Middle 80%"
      ),
      controversy_group = factor(controversy_group, levels = c("Top 10%", "Bottom 10%")) # set only for top and bottom 10%
    )

  # define bins 
  n_breaks <- c(0, 150, 300, Inf)
  n_labels <- c("≤150", "≤300", ">300")

  grouping_info <- temp2 |>
    group_by(label_specification, year, controversy_group) |>
    summarise(
      n = n(),
      .groups = "drop"
    ) |>
    mutate(
      n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
      n_group = factor(n_group, levels = n_labels)
    )

  n_values <- temp2 |>
    filter(!is.na(value_final)) |> # use only non-null observations (shown in plot)
    group_by(label_specification) |> # for all companies regardless of rating 
    summarise(n = n(), .groups = "drop")

  # summarise values
  p_summary <- temp2 |>
    group_by(label_specification, year, controversy_group) |> 
    summarise(
      p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Overall median
  overall_median <- temp2 |> 
    group_by(label_specification, year) |> 
    summarise(
      median = quantile(value_final, 0.50, na.rm = TRUE),
      .groups = "drop"
    ) |> 
    mutate(controversy_group = "All companies")

  p_summary_all <- bind_rows(p_summary, overall_median) # bind rows

  # set y axis limits independently for emission metrics
  emission_indicators <- c(
    "Total scope 1 emissions [tCO₂eq]",
    "Total scope 2 emissions (market-based) [tCO₂eq]",
    "Total scope 3 emissions [tCO₂eq]",
    "Category 6 Travel [tCO₂eq]"
  )

  shared_emission_max <- p_summary |>
    filter(label_specification %in% emission_indicators) |>
    summarise(ymax = max(p90, q3, na.rm = TRUE)) |>
    pull(ymax) * 1.1


  plotting <- function(indicator) {
    indicator_title <- switch(indicator,
      "Total scope 1 emissions [tCO₂eq]" = expression(Total~scope~1~emissions~"[tCO"[2]*"eq]"),
      "Total scope 2 emissions (market-based) [tCO₂eq]" = expression(Total~scope~2~emissions~"[tCO"[2]*"eq]"),
      "Total scope 3 emissions [tCO₂eq]" = expression(Total~Scope~3~emissions~"[tCO"[2]*"eq]"),
      "Scope 3 emissions (Cloud computing and data center services) [tCO₂eq]" = expression(Scope~3~emissions~(Cloud~computing~and~data~center~services)~"[tCO"[2]*"eq"),
      "Category 6 Travel [tCO₂eq]" = expression(atop(Scope~3~emissions, (Category~6~Travel)~"[tCO"[2]*"eq]")),
      "Scope 3 emissions (Commuting) [tCO₂eq]" = expression(Scope~3~emissions~(Commuting)~"[tCO"[2]*"eq"),
      "Energy consumption [MWh]" = expression(Energy~consumption~"[MWh]"),
      "Renewable sources in total energy consumption [%]" = expression(atop("Renewable sources in total", "energy consumption [%]")),
      "Fossil sources in total energy consumption [%]" = expression(atop("Fossil sources in total", "energy consumption [%]")),
      "Water consumption [m³]" = expression(Water~consumption~"[m"^3*"]"),
      "Total waste generated [t]" = expression(Total~waste~generated~"[t]"),
      "Non-recycled waste [%]" = expression("Non-recycled~waste [%]"),
      expression(.(indicator))
    )
    
  data_filtered <- p_summary_all |> # use p_summary_all
    filter(label_specification == indicator) |>
    left_join(grouping_info, by = c("label_specification", "year", "controversy_group"))

    n_label <- n_values |>
      filter(label_specification == indicator) |>
      pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

    y_max <- if (indicator %in% emission_indicators) {
      shared_emission_max
    } else {
      max(data_filtered$q3, data_filtered$median, na.rm = TRUE) * 1.1
    }
    
    p <- ggplot(data_filtered, aes(x = year))

    if (indicator == "Scope 3 emissions (Travel) [tCO₂eq]") {
        # add shaded area in the background
        p <- p +
          geom_rect(
            data = tibble(xmin = 2020, xmax = 2023, ymin = 1, ymax = 12000000),
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey85", alpha = 0.3, inherit.aes = FALSE
          ) +

          # add annotation text in the foreground
          annotate(
            "text",
            x = 2020.3, y = 500000,
            label = "COVID-19\npandemic",
            hjust = 0, size = 4, color = "grey30"
          )
    }

    p <- p + 
    geom_line(
        data = data_filtered |> filter(controversy_group == "All companies"),
        aes(y = median, color = controversy_group, linetype = controversy_group),
        linewidth = 1
      ) +
      geom_line(
        data = data_filtered |> filter(controversy_group %in% c("Top 10%", "Bottom 10%")), # show median for separate groups
        aes(y = median, color = controversy_group, linetype = controversy_group),
        size = 1
      ) 
    
      if (indicator %in% c(
      "Renewable sources in total energy consumption [%]",
      "Non-recycled waste [%]"
      )) {
        p <- p +
          scale_y_continuous(
            limits = c(
              min(data_filtered$p10, na.rm = TRUE) * 0.9,
              max(data_filtered$median, na.rm = TRUE) * 1.1
              ),
            labels = scales::comma_format()
            )
        } else {
          p <- p +
            scale_y_continuous(
              trans = "log10",
              labels = scales::comma_format(),
              limits = c(1, y_max),
              oob = scales::oob_keep
            )
        }
    
    p <- p +
      scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
      scale_color_manual(
        values = c("All companies" = "grey50", "Top 10%" = "#2E6F40", "Bottom 10%" = "#C94C4C"), # show only for top / bottom 10%
        name = NULL
      ) +
      scale_linetype_manual(
        values = c("All companies" = "dashed", "Top 10%" = "solid", "Bottom 10%" = "solid"), # show only for top / bottom 10%
        name = NULL
      ) +
    
      theme_minimal() +
      theme(
        plot.title = element_text(size = 14, hjust = 0.5),
        panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
        panel.grid.minor.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        axis.ticks.length = unit(0.15, "cm"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
      ) +
      labs(
        title = bquote(.(indicator_title)),
        x = "",
        y = parse(text = "")
      )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
        )

      p <- p + annotation_custom(label_grob)

      ggsave(
        filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = 'png'
      )

      return(p)
  }


  plots <- lapply(selected_indicators, plotting)

  pfinal <- wrap_plots(plots) +
    plot_layout(
      ncol = 3,
      guides = "collect"
    ) +
    plot_annotation(
      tag_levels = "a"
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 20),
      legend.position = "bottom",
      legend.box = "vertical",
      legend.justification = "center",
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 16)
    )

  ggsave(
      here("fig", "fig-env-impacts-controversy.pdf"),
      pfinal,
      width = 16, height = 15, dpi = 400, units = "in", device = cairo_pdf
  )

  knitr::include_graphics(here("fig/fig-env-impacts-controversy.pdf"))
}
```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Social performance top vs bottom 10% - ESG controversies score"`

```{r soc-controversies}
if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {
  #| fig-width: 16
  #| fig-height: 15

  selected_indicators <- c("Employee turnover [%]",
                          "Employees covered by collective bargaining agreements [%]",
                          "# Training hours per employee",
                          "# Days lost to work-related injuries, ill health, and fatalities among employees",
                          #"Employees entitled to take family-related leaves [%]",
                          "Female employees in top managment [%]",
                          "Gender pay gap [%]",
                          "Annual remuneration ratio [%]",
                          "# Complaints filed by own workforce",
                          "Amount of fines, penalties, and compensation for damages as result of incidents and complaints [USD]"
                          #"# Severe human rights incidents connected to own workforce"
                          )

  temp2 <- temp |>
    mutate(label_specification = recode(label_specification, !!!renamer_soc),
          label_specification = factor(label_specification, levels = unname(renamer_soc))
          ) |>
    mutate(
      ref_esg_controversies_score_n = recode(ref_esg_controversies_score_lagged, !!!controversy_score_map)
    ) |> 
    filter(label_specification %in% selected_indicators) |>
    select(firm, year, label_specification, value_final, ref_esg_controversies_score_n)

  # Assign groups based on controversy score percentiles (yearly)
  temp2 <- temp2 |>
    group_by(year) |> 
    mutate(
      top10 = ref_esg_controversies_score_n >= quantile(ref_esg_controversies_score_n, 0.90, na.rm = TRUE),
      bottom10 = ref_esg_controversies_score_n <= quantile(ref_esg_controversies_score_n, 0.10, na.rm = TRUE)
    ) |>
    ungroup() |>
    mutate(
      controversy_group = case_when(
        top10 ~ "Top 10%",
        bottom10 ~ "Bottom 10%",
        TRUE ~ "Middle 80%"
      ),
      controversy_group = factor(controversy_group, levels = c("Top 10%", "Bottom 10%"))
    )

  # define bins 
  n_breaks <- c(0, 150, 300, Inf)
  n_labels <- c("≤150", "≤300", ">300")

  grouping_info <- temp2 |>
    group_by(label_specification, year, controversy_group) |>
    summarise(
      n = n(),
      .groups = "drop"
    ) |>
    mutate(
      n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
      n_group = factor(n_group, levels = n_labels)
    )

  n_values <- temp2 |>
    filter(!is.na(value_final)) |> 
    group_by(label_specification) |>
    summarise(n = n(), .groups = "drop")

  # summarise values
  p_summary <- temp2 |>
    group_by(label_specification, year, controversy_group) |> 
    summarise(
      p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Overall median
  overall_median <- temp2 |> 
    group_by(label_specification, year) |> 
    summarise(
      median = quantile(value_final, 0.50, na.rm = TRUE),
      .groups = "drop"
    ) |> 
    mutate(controversy_group = "All companies")

  p_summary_all <- bind_rows(p_summary, overall_median) # bind rows

  plotting <- function(indicator) {

    data_filtered <- p_summary_all |>
    filter(label_specification == indicator) |>
    left_join(grouping_info, by = c("label_specification", "year", "controversy_group"))

    n_label <- n_values |>
      filter(label_specification == indicator) |>
      pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

    y_max <- if (indicator %in% emission_indicators) {
      shared_emission_max
    } else {
      max(data_filtered$q3, data_filtered$median, na.rm = TRUE) * 1.1
    }
    
    p <- ggplot(data_filtered, aes(x = year))

    p <- p +
      geom_line(
        data = data_filtered |> filter(controversy_group == "All companies"),
        aes(y = median, color = controversy_group, linetype = controversy_group),
        linewidth = 1
      ) +
      geom_line(
        data = data_filtered |> filter(controversy_group %in% c("Top 10%", "Bottom 10%")), # show median for separate groups
        aes(y = median, color = controversy_group, linetype = controversy_group),
        size = 1
      ) 
    
      if (indicator %in% c("Employee turnover [%]",
                          "Employees covered by collective bargaining agreements [%]",
                          "Female employees in top managment [%]",
                          "Gender pay gap [%]",
                          "Annual remuneration ratio [%]")) {
        p <- p + scale_y_continuous(
          limits = c(min(data_filtered$p10, na.rm = TRUE) * 0.9,
                    max(data_filtered$median, na.rm = TRUE) * 1.1),
          labels = scales::comma_format()
          )
      } else {
          p <- p + scale_y_continuous(
            trans = "log10",
            labels = scales::comma_format(),
            limits = c(1, y_max),
            oob = scales::oob_keep
            )
      }
    
      
    p <- p +
      scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
      scale_color_manual(
        values = c("All companies" = "grey50", "Top 10%" = "#2E6F40", "Bottom 10%" = "#C94C4C"),
        name = NULL
      ) +
      scale_linetype_manual(
        values = c("All companies" = "dashed", "Top 10%" = "solid", "Bottom 10%" = "solid"),
        name = NULL
      ) +

      theme_minimal() +
      theme(
        plot.title = element_text(size = 14, hjust = 0.5),
        panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
        panel.grid.minor.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        axis.ticks.length = unit(0.15, "cm"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
      ) +
      labs(
        title = stringr::str_wrap(indicator, width = 45),
        x = "",
        y = parse(text = "")
      )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
        )

      p <- p + annotation_custom(label_grob)

      ggsave(
        filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = 'png'
      )

      return(p)
  }


  plots <- lapply(selected_indicators, plotting)

  pfinal <- wrap_plots(plots) +
    plot_layout(
      ncol = 3,
      guides = "collect"
    ) +
    plot_annotation(
      tag_levels = "a"
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 20),
      legend.position = "bottom",
      legend.box = "vertical",
      legend.justification = "center",
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 16)
    )

  ggsave(
      here("fig", "fig-soc-impacts-controversy.pdf"),
      pfinal,
      width = 15, height = 15.5, dpi = 400, units = "in", device = cairo_pdf
  )

  knitr::include_graphics(here("fig/fig-soc-impacts-controversy.pdf"))
}
```

`r if (file.exists(file.path(datasets_dir, "firm_data.csv"))) "### Governance performance top vs bottom 10% - ESG controversies score"`

```{r}
if (file.exists(file.path(datasets_dir, "firm_data.csv"))) {
  #| fig-width: 15
  #| fig-height: 5.5


  selected_indicators <- c("Independent board members [%]",
                          #"# Confirmed incidents of corruption and bribery",
                          "Lobbying expenses [USD]",
                          #"Political contributions made [EUR]",
                          #"Variable remuneration dependent on sustainability-targets [%]"
                          "# Days to pay invoice"
                          )

  temp2 <- temp |>
    mutate(label_specification = recode(label_specification, !!!renamer_gov),
          label_specification = factor(label_specification, levels = unname(renamer_gov))
          ) |>
    mutate(
      ref_esg_controversies_score_n = recode(ref_esg_controversies_score_lagged, !!!controversy_score_map)
    ) |>
    filter(label_specification %in% selected_indicators) |>
    select(firm, year, label_specification, value_final, ref_esg_controversies_score_n)

  # Assign groups based on controversy score percentiles (yearly)
  temp2 <- temp2 |>
    group_by(year) |> 
    mutate(
      top10 = ref_esg_controversies_score_n >= quantile(ref_esg_controversies_score_n, 0.90, na.rm = TRUE),
      bottom10 = ref_esg_controversies_score_n <= quantile(ref_esg_controversies_score_n, 0.10, na.rm = TRUE)
    ) |>
    ungroup() |>
    mutate(
      controversy_group = case_when(
        top10 ~ "Top 10%",
        bottom10 ~ "Bottom 10%",
        TRUE ~ "Middle 80%"
      ),
      controversy_group = factor(controversy_group, levels = c("Top 10%", "Bottom 10%"))
    )

  # define bins 
  n_breaks <- c(0, 150, 300, Inf)
  n_labels <- c("≤150", "≤300", ">300")

  grouping_info <- temp2 |>
    group_by(label_specification, year, controversy_group) |>
    summarise(
      n = n(),
      .groups = "drop"
    ) |>
    mutate(
      n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
      n_group = factor(n_group, levels = n_labels)
    )

  n_values <- temp2 |>
    filter(!is.na(value_final)) |> 
    group_by(label_specification) |>
    summarise(n = n(), .groups = "drop")

  # summarise values
  p_summary <- temp2 |>
    group_by(label_specification, year, controversy_group) |> 
    summarise(
      p10 = max(quantile(value_final, 0.10, na.rm = TRUE), 1),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Overall median
  overall_median <- temp2 |> 
    group_by(label_specification, year) |> 
    summarise(
      median = quantile(value_final, 0.50, na.rm = TRUE),
      .groups = "drop"
    ) |> 
    mutate(controversy_group = "All companies")

  p_summary_all <- bind_rows(p_summary, overall_median) # bind rows

  plotting <- function(indicator) {

    data_filtered <- p_summary_all |>
    filter(label_specification == indicator) |>
    left_join(grouping_info, by = c("label_specification", "year", "controversy_group"))

    n_label <- n_values |>
      filter(label_specification == indicator) |>
      pull(n)

    label_n <- paste0("italic(n)==", n_label)

    # use dynamic y-axis limits
    y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE)
    y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

    y_max <- if (indicator %in% emission_indicators) {
      shared_emission_max
    } else {
      max(data_filtered$q3, data_filtered$median, na.rm = TRUE) * 1.1
    }
    
    p <- ggplot(data_filtered, aes(x = year)) +
      geom_line(
        data = data_filtered |> filter(controversy_group == "All companies"),
        aes(y = median, color = controversy_group, linetype = controversy_group),
        linewidth = 1
      ) +
      geom_line(
        data = data_filtered |> filter(controversy_group %in% c("Top 10%", "Bottom 10%")), # show median for separate groups
        aes(y = median, color = controversy_group, linetype = controversy_group),
        size = 1
      ) 
    
      if (indicator %in% c("Independent board members [%]")) {
        p <- p + scale_y_continuous(
          limits = c(min(data_filtered$p10, na.rm = TRUE) * 0.9,
                    max(data_filtered$median, na.rm = TRUE) * 1.1),
          labels = scales::comma_format()
          )
      } else {
          p <- p + scale_y_continuous(
            trans = "log10",
            labels = scales::comma_format(),
            limits = c(1, y_max),
            oob = scales::oob_keep
            )
      }
    
    p <- p +
      scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
      scale_color_manual(
        values = c("All companies" = "grey50", "Top 10%" = "#2E6F40", "Bottom 10%" = "#C94C4C"),
        name = NULL
      ) +
      scale_linetype_manual(
        values = c("All companies" = "dashed", "Top 10%" = "solid", "Bottom 10%" = "solid"),
        name = NULL
      ) +
      # scale_shape_manual(
      #   values = setNames(c(21, 21, 21), n_labels),
      #   name = "Reporting intensity (n)",
      #   guide = guide_legend(
      #     override.aes = list(
      #       fill = c("white", "#2E6F40", "#2E6F40"),
      #       size = c(3.5, 5.5, 5.5)
      #       )
      #     ),
      #   drop = FALSE
      #   ) +
      # scale_fill_manual(
      #   values = setNames(c("white", "#2E6F40", "#2E6F40"), n_labels),
      #   drop = FALSE
      # ) +
      # scale_size_manual(
      #   values = setNames(c(3.5, 3.5, 5.5), n_labels),
      #   drop = FALSE
      # ) +
      # guides(
      #   fill = "none",
      #   size = "none"
      # ) +

      theme_minimal() +
      theme(
        plot.title = element_text(size = 14, hjust = 0.5),
        panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
        panel.grid.minor.y = element_blank(),
        axis.ticks = element_line(color = "grey"),
        axis.ticks.length = unit(0.15, "cm"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
      ) +
      labs(
        title = stringr::str_wrap(indicator, width = 45),
        x = "",
        y = parse(text = "")
      )

    label_grob <- grid::textGrob(
        label = parse(text = label_n),
        x = unit(0.98, "npc"),
        y = unit(0.95, "npc"),
        just = c("right", "bottom"),
        gp = gpar(fontsize = 14)
        )

      p <- p + annotation_custom(label_grob)

      ggsave(
        filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
        plot = p,
        width = 5, height = 5, dpi = 400, units = "in", device = 'png'
      )

      return(p)
  }


  plots <- lapply(selected_indicators, plotting)

  pfinal <- wrap_plots(plots) +
    plot_layout(
      ncol = 3,
      guides = "collect"
    ) +
    plot_annotation(
      tag_levels = "a"
    ) &
    theme(
      plot.tag = element_text(face = "bold", size = 20),
      legend.position = "bottom",
      legend.box = "vertical",
      legend.justification = "center",
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 16)
    )

  ggsave(
      here("fig", "fig-gov-impacts-controversy.pdf"),
      pfinal,
      width = 15, height = 5.5, dpi = 400, units = "in", device = cairo_pdf
  )

  knitr::include_graphics(here("fig/fig-gov-impacts-controversy.pdf"))
}
```

### Scope 3 emission categories

```{r}
# Scope 3 emission categories
#| fig-width: 15
#| fig-height: 16

selected_indicators <- c(
  "Category 1 Purchased goods and services [tCO₂eq]",
  #"Category 1.1 Cloud computing and data center services [tCO₂eq]",
  "Category 2 Capital goods [tCO₂eq]",
  "Category 3 Fuel and energy-related activities [tCO₂eq]",
  "Category 4 Upstream transportation and distribution [tCO₂eq]",
  "Category 5 Waste generated in operations [tCO₂eq]",
  "Category 6 Travel [tCO₂eq]",
  "Category 7 Commuting [tCO₂eq]",
  "Category 8 Upstream leased assets [tCO₂eq]",
  "Category 9 Downstream transportation and distribution [tCO₂eq]",
  "Category 10 Processing of sold products [tCO₂eq]",
  "Category 11 Use of sold products [tCO₂eq]",
  "Category 12 End-of-life treatment of sold products [tCO₂eq]",
  "Category 13 Downstream leased assets [tCO₂eq]",
  "Category 14 Franchises [tCO₂eq]",
  "Category 15 Investments [tCO₂eq]"
  )

# define bins 
n_breaks <- c(0, 150, 300, Inf)
n_labels <- c("≤150", "≤300", ">300")

grouping_info <- temp |>
  mutate(label_specification = recode(label_specification, !!!renamer_env)) |>
  filter(topic == "env", label_specification %in% selected_indicators) |>
  drop_na(value_final) |>
  group_by(label_specification) |>
  mutate(total = n()) |>
  group_by(label_specification, year) |>
  summarise(
    n = n(),
    .groups = "drop"
  ) |> 
  mutate(
    n_group = cut(n, breaks = n_breaks, labels = n_labels, right = TRUE),
    n_group = factor(n_group, levels = n_labels)  # enforce consistent order
  )

p_env <- temp |> 
    select(firm, year, topic, label_specification, value_final) |> 
    mutate(label_specification = recode(label_specification, !!!renamer_env),
           label_specification = factor(label_specification, levels = unname(renamer_env))
           ) |>
    filter(topic == "env", label_specification %in% selected_indicators) |> 
    drop_na(value_final) |>
    group_by(label_specification, year) |> 
    ungroup() |> 
    group_by(label_specification, year) |> 
    summarise(
      p10 = quantile(value_final, 0.10, na.rm = TRUE),
      q1 = quantile(value_final, 0.25, na.rm = TRUE),
      median = quantile(value_final, 0.50, na.rm = TRUE),
      q3 = quantile(value_final, 0.75, na.rm = TRUE),
      p90 = quantile(value_final, 0.90, na.rm = TRUE), 
      .groups = "drop"
    ) |>
    left_join(grouping_info, by = c("label_specification", "year")) |>
    mutate(n_group = factor(n_group, levels = n_labels))

n_values <- temp |>
  mutate(label_specification = recode(label_specification, !!!renamer_env),
           label_specification = factor(label_specification, levels = unname(renamer_env))
           ) |>
  filter(topic == "env", label_specification %in% selected_indicators) |>
  drop_na(value_final) |>
  group_by(label_specification) |>
  summarise(n = n(), .groups = "drop")

# set y axis limits independently for emission metrics
emission_indicators <- selected_indicators

shared_emission_max <- p_env |>
  filter(label_specification %in% emission_indicators) |>
  summarise(ymax = max(p90, q3, na.rm = TRUE)) |>
  pull(ymax) * 1.1


plotting <- function(indicator) {
  indicator_title <- switch(indicator,
  "Category 1 Purchased goods and services [tCO₂eq]" = 
    expression(Purchased~goods~and~services~"[tCO"[2]*"eq]"),
  "Category 2 Capital goods [tCO₂eq]" = 
    expression(Capital~goods~"[tCO"[2]*"eq]"),
  "Category 3 Fuel and energy-related activities [tCO₂eq]" = 
    expression("Fuel and energy-related activities [tCO"[2]*"eq]"),
  "Category 4 Upstream transportation and distribution [tCO₂eq]" = 
    expression(Upstream~transportation~and~distribution~"[tCO"[2]*"eq]"),
  "Category 5 Waste generated in operations [tCO₂eq]" = 
    expression(Waste~generated~"in"~operations~"[tCO"[2]*"eq]"),
  "Category 6 Travel [tCO₂eq]" = 
    expression(Travel~"[tCO"[2]*"eq]"),
  "Category 7 Commuting [tCO₂eq]" = 
    expression(Commuting~"[tCO"[2]*"eq]"),
  "Category 8 Upstream leased assets [tCO₂eq]" = 
    expression(Upstream~leased~assets~"[tCO"[2]*"eq]"),
  "Category 9 Downstream transportation and distribution [tCO₂eq]" = 
    expression(Downstream~transportation~and~distribution~"[tCO"[2]*"eq]"),
  "Category 10 Processing of sold products [tCO₂eq]" = 
    expression(Processing~of~sold~products~"[tCO"[2]*"eq]"),
  "Category 11 Use of sold products [tCO₂eq]" = 
    expression(Use~of~sold~products~"[tCO"[2]*"eq]"),
  "Category 12 End-of-life treatment of sold products [tCO₂eq]" = 
    expression("End-of-life treatment of sold products [tCO"[2]*"eq]"),
  "Category 13 Downstream leased assets [tCO₂eq]" = 
    expression(Downstream~leased~assets~"[tCO"[2]*"eq]"),
  "Category 14 Franchises [tCO₂eq]" = 
    expression(Franchises~"[tCO"[2]*"eq]"),
  "Category 15 Investments [tCO₂eq]" = 
    expression(Investments~"[tCO"[2]*"eq]"),
  
  # fallback default
  expression(.(indicator))
  )

  data_filtered <- p_env |> filter(label_specification == indicator) |> 
    mutate(n_group = factor(n_group, levels = n_labels)) |> 
    complete(
      year,
      label_specification,
      n_group = factor(n_labels, levels = n_labels)
    )
  
  n_label <- n_values |>
    filter(label_specification == indicator) |>
    pull(n)
  
  label_n <- paste0("italic(n)==", n_label)
  
  # use dynamic y-axis limits
  y_min_raw <- min(data_filtered$p10, data_filtered$q1, na.rm = TRUE) 
  y_min <- if (y_min_raw < 1) 1 else 1 # set limits  <1 if needed

  y_max <- if (indicator %in% emission_indicators) {
    shared_emission_max
  } else {
    max(data_filtered$p90, data_filtered$q3, na.rm = TRUE) * 1.1
  }
  
  p <- ggplot(data_filtered, aes(x = year)) 
  
  if (indicator == "Scope 3 emissions (Travel) [tCO₂eq]") {
      # add shaded area in the background
      p <- p + 
        geom_rect(
          data = tibble(xmin = 2020, xmax = 2023, ymin = 1, ymax = 12000000),
          aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
          fill = "grey85", alpha = 0.3, inherit.aes = FALSE
        ) +
    
        # add annotation text in the foreground
        annotate(
          "text",
          x = 2020.3, y = 500000,
          label = "COVID-19\npandemic",
          hjust = 0, size = 4, color = "grey30"
        )
  }
  
  p <- p +
    geom_ribbon(data = filter(data_filtered, !is.na(q1) & !is.na(q3)),
                aes(ymin = q1, ymax = q3), fill = "#CFFCD8", alpha = 0.3) +
    geom_line(data = filter(data_filtered, !is.na(q1)),
              aes(y = q1, color = "25th"), linetype = "dashed") + 
    geom_line(data = filter(data_filtered, !is.na(median)),
              aes(y = median, color = "50th"), size = 1) + 
    geom_line(data = filter(data_filtered, !is.na(q3)),
              aes(y = q3, color = "75th"), linetype = "dashed") + 
    geom_point(aes(y = median, shape = n_group, size = n_group, fill = n_group),
               color = "#2E6F40") +
    geom_line(data = filter(data_filtered, !is.na(p10)),
              aes(y = p10, color = "10th"), linetype = "dashed", linewidth = 0.6) +
    geom_line(data = filter(data_filtered, !is.na(p90)),
              aes(y = p90, color = "90th"), linetype = "dashed", linewidth = 0.6) +
    
    scale_y_continuous(
      trans = "log10",
      labels = scales::comma_format(),
      limits = c(1, y_max),
      oob = scales::oob_keep
    ) +

    scale_x_continuous(breaks = seq(2015, 2023, by = 2)) +
    scale_color_manual(values = c(
      "10th" = "#55C173", 
      "25th" = "#419759", 
      "50th" = "#2E6F40", 
      "75th" = "#419759", 
      "90th" = "#55C173"
    ), name = "Percentiles") +
    scale_shape_manual(
      values = setNames(c(21, 21, 21), n_labels),
      name = expression(bold("Reporting intensity (" * italic(n) * ")")),
      guide = guide_legend(
        override.aes = list(
          fill = c("white", "#2E6F40", "#2E6F40"),
          size = c(3.5, 5.5, 5.5)
          )
        ),
      drop = FALSE
      ) +
    scale_fill_manual(
      values = setNames(c("white", "#2E6F40", "#2E6F40"), n_labels),
      drop = FALSE
    ) +
    scale_size_manual(
      values = setNames(c(3.5, 3.5, 5.5), n_labels),
      drop = FALSE
    ) +
    guides(
      fill = "none",
      size = "none"
    ) +
    
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      panel.border = element_rect(color = "grey", fill = NA, linewidth = 0.8),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
      panel.grid.minor.y = element_blank(),
      axis.ticks = element_line(color = "grey"),
      axis.ticks.length = unit(0.15, "cm"),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_text(size = 16)
    ) +
    labs(
      title = bquote(.(indicator_title)),
      x = "",
      y = parse(text = "")
    )
  
  label_grob <- grid::textGrob(
      label = parse(text = label_n),
      x = unit(0.98, "npc"),
      y = unit(0.95, "npc"),
      just = c("right", "bottom"),
      gp = gpar(fontsize = 14)
      )
    
    p <- p + annotation_custom(label_grob)
  
    ggsave(
      filename = here("fig", paste0("fig-", gsub("[^A-Za-z0-9]", "_", indicator), ".png")),
      plot = p,
      width = 5, height = 5, dpi = 400, units = "in", device = 'png'
    )
    
    return(p)
}


plots <- lapply(selected_indicators, plotting)

pfinal <- wrap_plots(plots) +
  plot_layout(
    ncol = 3,
    guides = "collect"
  ) +
  plot_annotation(
    tag_levels = "a"
  ) &
  theme(
    plot.tag = element_text(face = "bold", size = 20),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.justification = "center",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 16)
  )

ggsave(
    here("fig", "fig-env-impacts-scope3.pdf"),
    pfinal,
    width = 16, height = 25, dpi = 400, units = "in", device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-env-impacts-scope3.pdf"))

```

### Scope 3 emission categories per year

```{r}
selected_indicators <- c(
  "Category 1 Purchased goods and services [tCO₂eq]",
  "Category 2 Capital goods [tCO₂eq]",
  "Category 3 Fuel and energy-related activities [tCO₂eq]",
  "Category 4 Upstream transportation and distribution [tCO₂eq]",
  "Category 5 Waste generated in operations [tCO₂eq]",
  "Category 6 Travel [tCO₂eq]",
  "Category 7 Commuting [tCO₂eq]",
  "Category 8 Upstream leased assets [tCO₂eq]",
  "Category 9 Downstream transportation and distribution [tCO₂eq]",
  "Category 10 Processing of sold products [tCO₂eq]",
  "Category 11 Use of sold products [tCO₂eq]",
  "Category 12 End-of-life treatment of sold products [tCO₂eq]",
  "Category 13 Downstream leased assets [tCO₂eq]",
  "Category 14 Franchises [tCO₂eq]",
  "Category 15 Investments [tCO₂eq]"
  )

label_df <- tibble(
  full_label = selected_indicators,
  short_label = str_remove(selected_indicators, "\\s*\\[.*?\\]"),
  cat_number = as.numeric(str_extract(selected_indicators, "(?<=Category )\\d+(\\.\\d+)?"))
) |> 
  arrange(cat_number)

plot_scope3_n <- grouping_info |> 
  filter(label_specification %in% selected_indicators) |> 
  mutate(
    short_label = str_remove(label_specification, "\\s*\\[.*?\\]"),
    short_label = factor(short_label, levels = label_df$short_label)
  ) |> 
  ggplot(aes(x = year, y = n)) +
  geom_line(aes(color = short_label), linewidth = 1) +
  geom_point(aes(color = short_label), size = 2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_x_continuous(breaks = seq(min(grouping_info$year), max(grouping_info$year), by = 1)) +
  labs(
    title = "",
    x = NULL,
    y = NULL,
    color = "Indicator"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dotted", color = "gray60"),
    panel.grid.minor.y = element_blank(),
    legend.position = "right",
    legend.direction =  "vertical",
    legend.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


# Optional: save to file
ggsave(
  here("fig", "fig-scope3-n-per-year.pdf"),
  plot_scope3_n,
  width = 20, height = 10, dpi = 400, device = cairo_pdf
)

knitr::include_graphics(here("fig/fig-scope3-n-per-year.pdf"))

```


\newpage
